---
title: "Nigeria chloropleth prioritization index code"
author: "W.E., A.Y., H.H."
date: "3/28/2021"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Clear workspace

# Task: 

## 1) Create workspace to pull from git or pull from the same folder as Rproject

Remove objects from current workspace, Clear startup/console, close plots

```{r}
# Remove all objects from the current workspace (R memory) --------------------
rm(list = ls())

# This function closes the specified plot (by default the current device) -----
# and if it is an imguR device, uploads the plots for web hosting -------------
dev.off()

# Clear startup screen/console in R / RStudio ----------------------------------
cat("\014") 

# Set a 
```

## Explore color palette

```{r}
# Explore color palette ---
# palette_explorer()

```


## Load packages

## Task: 
## 1) filter through these and determine which packages you don't need
## 2) determine which packages might interfere with concurrently loaded packages
##    and make a note of it

```{r}

# Load packages - these are the main packages utilized
library(raster)
library(sp)
library(tmap)
library(tmaptools)
library(BAMMtools)
library(exactextractr)

# Remaining packages that may be utilized

library(readxl)
library(ggplot2)
library(malariaAtlas)
library(exactextractr)
library(maptools)
library(sf)
# library("spDataLarge")
library(dplyr)
library(GADMTools)

library(maptools)
library(readr)
library(foreign)   ; library(tsModel) ; library("lmtest") ; library("Epi")
library("splines") ; library("vcd")
library(reshape2)  ; library(hablar)
library(tidyr)     
library (viridis)
library(data.table)
library(forecast)  ; library(MASS)
library(tseries)   ; library(scales)
library(tsModel)   ; library(extrafont)
library(lmtest)    ; library(tidyverse)
library(stargazer) ; library(RColorBrewer)
library(readxl)    ; library(olsrr)
library(Hmisc)
library(MASS)
library(dplyr)
library(devEMF)
library(padr)
library(zoo)
library(tidyverse)
library(naniar)
library(GGally)
library(cartogram)
library(mgcv)
library(rgdal)
library(leaflet)
library(arsenal)
require(data.table)
require(RCurl)
require(R.utils)
require(gdalUtils)
require(parallel)
```

## Load necessary administrative boundary shapefiles and plot

```{r}
# -----------------------------------------------------------------------------
# Nigeria new chloropleth map code
# Author: Will Eaton

# Author comment
# File created for Nigerian data management and creation of chloropleth maps
# for 2020 ITN prioritzation within local government areas (LGAs) of Nigeria

# Inputs include:
#   1) master data set (compiled by A. Young)
#   2) spatial interpolation outputs (created by M. Worges)

# Outputs include: 
#   1) weighted malaria risk / ITN priortiziation map
#   2) unweighted malaria risk / ITN priortization map
#   3) individual indicator input maps for creation of final malaria risk/ ITN
#      prioritization index

# Date created: 7-12-20
# Date modified: 8-22-20
# Local drive location: /Volumes/Samsung T5/LACIE MacOS Extended/Tulane Research Projects/Malaria Consortium/Methodology/R Code/MC R Code/Malaria-Consortium
# git: https://github.com/weaton5061/Malaria-Consortium.git
# -----------------------------------------------------------------------------
#-------------------------------------------------------------------------------
# TASKS:
# 1) Where is differentiation with n = 774 and n = 775 LGAs /shape
#    files
#-------------------------------------------------------------------------------

# Code color palette modifications --------------------------------------------
# col = RColorBrewer::brewer.pal(4, "YlOrRd"),

# Set (modifiable) working directory ------------------------------------------
# setwd("D:\\LACIE MacOS Extended\\Tulane Research Projects\\Malaria Consortium\\Nigeria Data\\Maps\\smc") # PC

# Get admin boundaries ----------------------------------------------------
# GADM, the Database of Global Administrative Areas, is a high-resolution 
# database of country administrative areas, with a goal of “all countries, 
# at all levels, at any time period.
adm0.nga <- subset(getData("GADM", country = "NGA", level = 0)) # Country
# plot(adm0.nga)
adm1.nga <- subset(getData("GADM", country = "NGA", level = 1)) # n = 37 states
# has n = 37 unique identifiers using NAME_1 and GID_1
# plot(adm1.nga)

# Bring in HDX shapefile for comparison (n=75 level 1) ------------------------
# nga_ocha_level_1 <- st_read(dsn = "D:\\LACIE MacOS Extended\\Tulane Research Projects\\Malaria Consortium\\Nigeria Data\\Shapefiles\\nga_adm_osgof_20190417\\nga_admbnda_adm0_osgof_20190417.shp") # PC

# adm2.nga appears to be LGAs with unique identifiers called GID_2 ------------
# appears to have n = 775 observations for GID_2
# Note: If I use NAME_2, only n = 770 unique identifiers exist
adm2.nga <- subset(getData("GADM", country = "NGA", level = 2)) # LGAs
# plot(adm2.nga)

# Plot map in tmap ------------------------------------------------------------
tm_shape(adm2.nga) + tm_polygons("GID_2") + tm_legend(show=FALSE)

# View adm2.nga as data frame -------------------------------
adm2.nga.df <- as.data.frame(adm2.nga)
View(adm2.nga.df)                       # View adm2.nga.df

```

### Make nga states map n = 75 states

```{r}
# --------------------------------------------------------------------------
# Make nga states map n = 75 states (text size relative to polygon area) ---
# --------------------------------------------------------------------------
tm_shape(adm1.nga) + tm_fill("GID_1", palette = "PiYG") + tm_borders("NAME_1")

# use for state borders below 
# tm_shape(adm1.nga)  + tm_borders(col = "grey40", lwd = 1)

# Save nga states map to object --------------------------------------------
# Note: not working on Will's mac 8-23-20, however, maps saved appropriately
# to working directory listed below
nga_states_map <- tm_shape(adm1.nga) + 
    tm_fill(col = "MAP_COLORS", legend.show = FALSE) + 
    tm_layout(frame = FALSE) + 
    tm_polygons(col = "MAP_COLORS", legend.show = FALSE) + 
    tm_borders(col = "grey40", lwd = 1) + tm_text("NAME_1", size = "AREA") + 
    tm_scale_bar() + tm_compass( position = c("right","top"))

# Save nga states (with 0.5 size text) to map object
nga_states_map_0_5_text <- tm_shape(adm1.nga) + 
    tm_fill(col = "MAP_COLORS", legend.show = FALSE) + 
    tm_layout(frame = FALSE) + 
    tm_polygons(col = "MAP_COLORS", legend.show = FALSE) + 
    tm_borders(col = "grey40", lwd = 1) + tm_text("NAME_1", size = 0.5) + 
    tm_scale_bar() + tm_compass( position = c("right","top"))

# Save nga states (with 0.75 size text) to map object
nga_states_map_0_75_text <- tm_shape(adm1.nga) + 
    tm_fill(col = "MAP_COLORS", legend.show = FALSE) + 
    tm_layout(frame = FALSE) + 
    tm_polygons(col = "MAP_COLORS", legend.show = FALSE) + 
    tm_borders(col = "grey40", lwd = 1) + tm_text("NAME_1", size = 0.75) + 
    tm_scale_bar() + tm_compass( position = c("right","top"))

# Save nga states map ---------------------------------------------------------
# setwd("D:\\LACIE MacOS Extended\\Tulane Research Projects\\Malaria Consortium\\Nigeria Data\\Maps\\nga states") # Travis PC
# Maps saved appropriately to working directory
setwd("/Volumes/Samsung T5/LACIE MacOS Extended/Tulane Research Projects/Malaria Consortium/Nigeria Data/Maps/nga states") # Mac
tmap_save(nga_states_map, "nga_states_map.png",
          width =6.78, height=5, units ='in', asp = 0)
tmap_save(nga_states_map_0_5_text, "nga_states_map_0_5_text.png",
          width =6.78, height=5, units ='in', asp = 0)
tmap_save(nga_states_map_0_75_text, "nga_states_map_0_75_text.png",
          width =6.78, height=5, units ='in', asp = 0)

# Bring in OCHA Nigeria Shapefile with hopefully country boundary for water (Verify) ----------
hdx_nga_admin_0 <- st_read(dsn = "D:\\LACIE MacOS Extended\\Tulane Research Projects\\Malaria Consortium\\Nigeria Data\\Shapefiles\\nga_adm_osgof_20190417\\nga_admbnda_adm0_osgof_20190417.shp")
tm_shape(hdx_nga_admin_0) + tm_polygons("ADM0_EN") + tm_legend(show=FALSE)  + 
    tm_layout(frame = FALSE, legend.outside = TRUE, 
              legend.outside.position = "right")
# Verify shapefile - LOOKS GOOD!

# Bring in OCHA admin 2 shapefile for comparison -------------------------------------
# Appears to have n = 774 observations for ADM2_EN
hdx_nga_admin_2 <- st_read(dsn = "D:\\LACIE MacOS Extended\\Tulane Research Projects\\Malaria Consortium\\Nigeria Data\\Shapefiles\\nga_adm_osgof_20190417\\nga_admbnda_adm2_osgof_20190417.shp")
tm_shape(hdx_nga_admin_2) + tm_polygons("ADM2_EN") + tm_legend(show=FALSE) +
tm_layout(frame = FALSE, legend.outside = TRUE, 
          legend.outside.position = "right")
# Verify shapefile - LOOKS GOOD!
```

## Import data and peform data managment steps
### (!!!consider reorganizing and renaming this section!!!)

```{r}
# Bring in Nigeria current master dataset NGA_master6.csv ---------------------------------
# nga_master6 <- read.csv("//Volumes/Samsung T5/LACIE MacOS Extended/Tulane Research Projects/Malaria Consortium/Nigeria Data/NGA_master6_2.csv", header = TRUE) # Mac
# nga_master6 <- read.csv("D:\\LACIE MacOS Extended\\Tulane Research Projects\\Malaria Consortium\\Nigeria Data\\NGA_master6_2.csv", header = TRUE) #PC
# Bring in NGA_master6_2.xls excel file in as .xls instead of .csv
# See if this resolves variable renaming issues (NAME_2)
# setwd("D:\\LACIE MacOS Extended\\Tulane Research Projects\\Malaria Consortium\\Nigeria Data")
# setwd("/Users/willeaton/Box/Will External Drives/LACIE MacOS Extended/Malaria Consortium/Nigeria Data")
# nga_master6 <- read_excel("NGA_master6_2.xls")
nga_master6 <- read_excel("/Users/willeaton/Box/Will External Drives/LACIE MacOS Extended/Malaria Consortium/Nigeria Data/NGA_master6_2.xls")
# View(nga_master6)

# Bring in spatial interpolation data from M. Worges 
# (n = 775 LGAs, usingunique identifier = GID_2)
# setwd("D:\\LACIE MacOS Extended\\Tulane Research Projects\\Malaria Consortium\\Nigeria Data\\MWorges INLA SPDE") # Travis PC
# setwd("/Volumes/Samsung T5/LACIE MacOS Extended/Tulane Research Projects/Malaria Consortium/Nigeria Data/MWorges INLA SPDE")
# setwd("/Users/willeaton/Box/Will External Drives/LACIE MacOS Extended/Malaria Consortium/Nigeria Data/MWorges INLA SPDE")
# sp_interp_data <- read_excel("ZonalStats_NGA_V5.xlsx")
sp_interp_data <- read_excel("/Users/willeaton/Box/Will External Drives/LACIE MacOS Extended/Malaria Consortium/Nigeria Data/MWorges INLA SPDE/ZonalStats_NGA_V5.xlsx")

# Check sp_interp_data for missing values ------------------------------------
table(sp_interp_data$NetAvlblty.wghted, useNA = "always") # no missing values

# Note: net availability variable is the proportion of households with 
# at least 1 net per 2 people. 
# Will need to take complement value, create, class, 
# apply weight then, create score

# Create jenks/breaks --------------------------------------------
# With natural breaks classification (Jenks) Natural Breaks Jenks, 
# classes are based on natural groupings inherent in the data. 
# Class breaks are identified that best group similar values and 
# that maximize the differences between classes. The features are 
# divided into classes whose boundaries are set where there are 
# relatively big differences in the data values.
# 
# Natural breaks are data-specific classifications and not useful for 
# comparing multiple maps built from different underlying information.

###############################################
# Left off here 7-13-20 @ 11:54 pm
# TASK: 1) Figure out when to apply weight
# and 2) make net availability map
# 3) merge with master dataset again by GID_2
###############################################

```

## Net availability
## Jenks natural breaks steps (!!!rename this section!!!)

```{r}

# Net availability -----------------------------------------------
getJenksBreaks(sp_interp_data$NetAvlblty.wghted, 5, subset = NULL)  
# [1] 0.009014783 0.233290972 0.354107688 0.493038245 0.792025051
# (to use in legend)

# Create weighted complement net availability score variable ------------------
# variable in Matt's sp_interp_data is called NetAvlblty.wghted
sp_interp_data$net_avlblty_complement <- (1 - sp_interp_data$NetAvlblty.wghted)
# sp_interp_data$net_avlblty_score_complement_wghtd <- sp_interp_data$net_avlblty_complement * sp_interp_data$net_avlblty_wt

# Net availability complement jenks -----------------------------------------------
getJenksBreaks(sp_interp_data$net_avlblty_complement, 5, subset = NULL)  
# [1] 0.2079749 0.4951487 0.6451532 0.7655778 0.9909852

```

## Create net availability complement classes (!!! rename this section)

```{r}

# Assign net availability classes (numeric)
sp_interp_data$net_avlblty_complement_class[sp_interp_data$net_avlblty_complement > 0.7655778] <- 4     # classify as rank 4
sp_interp_data$net_avlblty_complement_class[sp_interp_data$net_avlblty_complement <= 0.7655778 & sp_interp_data$net_avlblty_complement > 0.6451532 ] <- 3     # classify as rank 3
sp_interp_data$net_avlblty_complement_class[sp_interp_data$net_avlblty_complement <= 0.6451532 & sp_interp_data$net_avlblty_complement > 0.4951487 ] <- 2     # classify as rank 2
sp_interp_data$net_avlblty_complement_class[sp_interp_data$net_avlblty_complement < 0.4951487] <- 1 # classify as rank 1

# Show the distribution of net availability complement class
table(sp_interp_data$net_avlblty_complement_class, useNA = "always")
#  1    2    3    4 <NA> 
# 78  220  248  229    0 

# Set net availability weight to 0.75
sp_interp_data$net_avlblty_wt <- 0.75

# Create net availability score variable --------------------------------------
sp_interp_data$weighted_net_avlblty_complement_score <- sp_interp_data$net_avlblty_wt * sp_interp_data$net_avlblty_complement_class

# Show distribution of weighted net availability complement score -------------
table(sp_interp_data$weighted_net_avlblty_complement_score, useNA = "always")
#   0.75  1.5 2.25    3 <NA> 
#     78  220  248  229    0 

# slim sp_interp_data to merge with nga_master6 ------------------------------
myvars <- c(6,15:26)
sp_interp_data_slim <- sp_interp_data[myvars]

# merge sp_interp_data_slim (13 variables) and ------------------------------
# nga_master6 datasets (103 variables) 
nga_master6_sp_combo <- nga_master6
nga_master6_sp_combo <- merge(x=nga_master6, y=sp_interp_data_slim, by="GID_2", all = TRUE)

# slim nga_master6_sp_combo dataset ------------------------------------------
# myvars2 <- c()
# Shorten name of master data set
nga_7 <- nga_master6_sp_combo

```

## Import COVID data and manage COVID data

```{r}
# 8-4-20 Bring in newest COVID-19 data for Nigeria
# Note: Newer COVID-19 data has been imported as of 8-27-20
nga_7_df_v2_covid <- read.csv("/Volumes/Samsung T5/LACIE MacOS Extended/Tulane Research Projects/Malaria Consortium/Nigeria Data/nga_7_df_v2_covid.csv")

#------------------------------------------------------------------------------
# 8-28-20 modification --------------------------------------------------------
#------------------------------------------------------------------------------
# 1) Bring in Alyssa's revised data set that has updated COVID-19 case count data
# 2) Dbl check yrs since last LLIN variable that have 2020 entries
nga_8_28_20.df <- read.csv("/Volumes/Samsung T5/LACIE MacOS Extended/Tulane Research Projects/Malaria Consortium/Nigeria Data/Datasets/nga_8_df_28_20.csv") # Mac

# Slim down nga_8_28_20.df data set
nga_8_28_20_vars <- c(2, 6, 15:22, 32:41, 51:92)
nga_8_28_20_slim.df <- nga_8_28_20.df[nga_8_28_20_vars]
View(nga_8_28_20_slim.df)

# Slim down nga_8_28_20_slim.df to just COVID variables
covid_8_28_20_vars <- c(1, 2, 16, 24, 25, 43, 56)
covid_8_28_20.df <- nga_8_28_20_slim.df[covid_8_28_20_vars]
View(covid_8_28_20.df)

# COVID jenks ----------------------------------------------------------------
# Obtain 5 classes of breaks for COVID 
# Updated jenks (8-28-20) for COVID cases
getJenksBreaks(covid_8_28_20.df$Covid_case_28_8_20, 5, subset = NULL)
# [1]     5  1142  2555  5094 18056

summary(covid_8_28_20.df$Covid_case_28_8_20)
# Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#   5     241     740    1403    1722   18056 

covid_8_28_20.df$COVID_class_aug_28[covid_8_28_20.df$Covid_case_28_8_20 > 5094] <- 4                                               # classify as rank (class) 4
covid_8_28_20.df$COVID_class_aug_28[covid_8_28_20.df$Covid_case_28_8_20 <= 5094 & covid_8_28_20.df$Covid_case_28_8_20 > 2555] <- 3 # classify as rank (class) 3
covid_8_28_20.df$COVID_class_aug_28[covid_8_28_20.df$Covid_case_28_8_20 <= 2555 & covid_8_28_20.df$Covid_case_28_8_20 > 1142] <- 2 # classify as rank (class) 2
covid_8_28_20.df$COVID_class_aug_28[covid_8_28_20.df$Covid_case_28_8_20 <= 1142] <- 1                                              # classify as rank (class) 1

# Assign covid weight
covid_8_28_20.df$covid_wt <- 0.28

# Create covid score variable (from Aug 2020 data)
covid_8_28_20.df$COVID_case_score_aug_8_wtd <- covid_8_28_20.df$covid_wt * covid_8_28_20.df$COVID_class_aug_28 

# copy weighted case score to previous weighted case score variable
covid_8_28_20.df$covid_case_score_wtd <- covid_8_28_20.df$COVID_case_score_aug_8_wtd

# View dataset for quick check
View(covid_8_28_20.df)

# How many different score categories? Are any missing? -----
table(covid_8_28_20.df$COVID_case_score_aug_8_wtd, useNA = "always")
# Previous score categories using July 10 case counts ----
# 0.28 0.56 0.84 1.12 <NA> 
#  361  225  169   20    0

# New score categories using July 21 case counts ----
# 0.28 0.56 0.84 1.12 <NA> 
#  545  153   57   20    0 

# Newest score categories using Aug 8 case counts ----
# 0.28 0.56 0.84 1.12 <NA> 
#  528  188   39   20    0 

# Convert covid_case_score_wtd to factor variable
covid_8_28_20.df$covid_case_score_wtd <- as.factor(covid_8_28_20.df$covid_case_score_wtd)

# Rename gid_2 to GID_2
covid_8_28_20.df$GID_2 <- covid_8_28_20.df$gid_2

```

## Malaria atlas data manipulation

```{r}
#------------------------------
# MALARIA ATLAS PROJECT DATA
#------------------------------

# Temperature suitability ----

# Show malaria atlas project data available----------
listData('raster', printed = TRUE)
listRaster(printed = TRUE)
#  Plasmodium falciparum Temperature Suitability 

# Download malaria atlas project temp suitability raster layer ---------------
NGA_shp <- getShp(ISO = "NGA", admin_level = "admin0")

# (From https://rdrr.io/cran/malariaAtlas/man/getRaster.html)
nga_temp_suit_raster <- getRaster(surface = "Plasmodium falciparum Temperature Suitability", shp = NGA_shp)
# Visualize raster in plot
autoplot_MAPraster(nga_temp_suit_raster) #THIS WORKS!!!

#----------------------------------------------------------------------------
# Temperature suitability index for P. falciparum transmission, 2010
#----------------------------------------------------------------------------
# Source: Gething, P. W., Van Boeckel, T. P., Smith, D. L., Guerra, C. A., 
# Patil, A. P., Snow, R. W., & Hay, S. I. (2011). 
# Modelling the global constraints of temperature on transmission of Plasmodium
# falciparum and P. vivax. Parasites and Vectors, 4(1), 92. 
# https://doi.org/10.1186/1756-3305-4-92
#----------------------------------------------------------------------------
# The normalized Z(T) index of temperature suitability that incorporates not
# just the duration but also the degree of suitability across an average year. 
# See text for a full explanation of all three metrics

# Assign temp suitability map to object (grey50 border color and grey50 tm_borders)
# This provides visualization of raster layer w/o LGA boundaries
temp_suit_raster_map <- tm_shape(nga_temp_suit_raster) +
    tm_raster("Plasmodium.falciparum.Temperature.Suitability", 
              # palette = pal8, 
              title = "Plasmodium falciparum Temperature Suitability") +
    tm_shape(adm2.nga) + tm_polygons("GID_2", lwd = 0.25, col = "gray3",
                                     border.col = "grey50") 

# This code provides visualization of raster layer w/ LGA boundaries
temp_suit_raster_lga_map <- tm_shape(nga_temp_suit_raster) +
    tm_raster("Plasmodium.falciparum.Temperature.Suitability", 
              title = "Plasmodium falciparum Temperature Suitability") +
    tm_shape(adm2.nga) + tm_borders(col = "gray3") +
    tm_layout(frame = FALSE)

# set working directory for temp suitability map 
setwd("/Volumes/Samsung T5/LACIE MacOS Extended/Tulane Research Projects/Malaria Consortium/Nigeria Data/Maps/temp suitability") # Mac
# save temp suitability map to working directory listed above
tmap_save(temp_suit_raster_lga_map, "temp_suit_raster_lga_map.png",
          width =6.78, height=5, units ='in', asp = 0)

# Zonal statisitics LGAs & temperature suitability index ----------------
# Extract temperature suitability to LGA (n = 775)
adm2.nga.sf <- st_as_sf(adm2.nga)
# nga_temp_suit_extract_lga$mean_temp_suit <- extract(nga_temp_suit_raster, 
#                                                     adm2.nga.sf, fun=mean, 
#                                                     na.rm=TRUE, df=TRUE)
nga_temp_suit_extract_lga$mean_temp_suit <- exact_extract(nga_temp_suit_raster, adm2.nga.sf, 'mean') # this code did not work
# The code above did not work due to the following error:
# Error in (function (classes, fdef, mtable)  : 
# unable to find an inherited method for function ‘exact_extract’ for signature
# ‘"RasterLayer", "SpatialPolygonsDataFrame"’

# Convert nga_temp_suit_extract_lga to data frame for quick visual inspection --
nga_temp_suit_extract_lga.df <- as.data.frame(nga_temp_suit_extract_lga)
View(nga_temp_suit_extract_lga.df)

# slim down nga_temp_suit_extract_lga
nga_temp_suit_extract_lga_slim <- nga_temp_suit_extract_lga[c(6, 14, 15) ]
# convert to df
nga_temp_suit_extract_lga_slim.df <- as.data.frame(nga_temp_suit_extract_lga_slim)
# slim nga_temp_suit_extract_lga_slim.df data frame even more
temp_suit.df <- nga_temp_suit_extract_lga_slim.df[c(1,3)]

# Create new variable in nga_7 ------------------------------------------
# How many LGAs are receiving SMC? n = 72
# table(nga_master6$SMC_class)
#       0    1 <NA> 
#     703   72    0 

```

## SMC coverage

```{r}

# Create new SMC coverage weight = 0.82
nga_7$smc_weight_0_82 <- 0.82
# nga_7$smc_score_weighted <- nga_7$smc_weight_0_82 * nga_7$SMC_class calculated incorrectly

# States that received SMC include:
# Borno         6% coverage
# Yobe         27% coverage
# Katsina      26% coverage
# Jigawa      107% coverage
# Sokoto      111% coverage
# Zamfara     114% coverage

```

## Built pop

```{r}
# Manipulate built environment up presence layer  ------------------------
# from Landsat image collections
summary(nga_7$BuiltPop_2014)    
#    Min.  1st Qu.   Median     Mean  3rd Qu.     Max. 
# 0.00058  0.17165  0.60954  5.16307  1.99653 98.60358 

# built environment
getJenksBreaks(nga_7$BuiltPop_2014, 5, subset = NULL) # obtain 5 classes of breaks for Built_switch 
# [1] 5.822306e-04 6.389496e+00 2.261250e+01 5.156892e+01 9.860358e+01

# Create Built environ (rural/urban) weight variable -------
nga_7$Built_wt <- 0.54

# Create equal weight variable for data set ---------
nga_7$eq_wt <- 0.50

# Make BuiltPop_2014 variable a value between 0 and 1
nga_7$BuiltPop_hundredth <- (nga_7$BuiltPop_2014) / 100

# Create complement variable (1 - BuiltPop_hundredth value)
nga_7$Built_switch <- 1- nga_7$BuiltPop_hundredth 

# Create jenks / breaks # built environment ----------------------------------
# Obtain 5 classes of breaks for Built_switch
# For built switch, value near zero will represent low risk malaria, likely,
# much built area (urban), while value near 1 will represent higher risk malaria with
# less built area (rural)
getJenksBreaks(nga_7$Built_switch, 5, subset = NULL)  
# [1] 0.01396416 0.39074421 0.75832115 0.93383096 0.99999418

```

## Create built pop classes

```{r}


# Create built classes
nga_7$Built_class[nga_7$Built_switch >  0.93383096] <- 4                                          # classify as rank 4
nga_7$Built_class[nga_7$Built_switch <= 0.93383096   & nga_7$Built_switch > 0.75832115 ] <- 3     # classify as rank 3
nga_7$Built_class[nga_7$Built_switch <= 0.75832115   & nga_7$Built_switch > 0.39074421 ] <- 2     # classify as rank 2
nga_7$Built_class[nga_7$Built_switch <  0.39074421] <- 1                                          # classify as rank 1

# Create slim subset to inspect BuiltPop_2014
built_vars <- c(1,2,106, 116:118)
built_vars_slim <- nga_7[built_vars]
View(built_vars_slim)

# Create built class variable
nga_7$Built_class <- as.numeric(nga_7$Built_class) # convert Built_class class to numeric

# Built score calculation
nga_7$Built_score <- nga_7$Built_class * nga_7$Built_wt # create Built weighted score 

# Equal wt Built score calculation
nga_7$Built_eq_wt_score <- nga_7$Built_class * nga_7$eq_wt        #create built eq wt score

```

## Precipitation - rainfall

```{r}
# Remove precipitation months before June 2019
nga_7_subset <- subset(nga_7, select = -c(35:75, 87:98))

# Previous rainfall code that included negative values. Suspect negative values weren't converted
# to zeros prior to extraction and mean calculations
# Create mean annual rainfall (June 2019 - May 2020, the past 12 
# months of data)variable (by LGA)
# nga_7_subset$mean_an_rf <- ((nga_7_subset$mean_jun19_prec + nga_7_subset$mean_jul19_prec +
#     nga_7_subset$mean_aug19_prec + nga_7_subset$mean_sep19_prec + nga_7_subset$mean_oct19_prec +
#     nga_7_subset$mean_nov19_prec + nga_7_subset$mean_dec19_prec + nga_7_subset$mean_jan20_prec +
#     nga_7_subset$mean_feb20_prec + nga_7_subset$mean_mar20_prec + nga_7_subset$mean_apr20_prec +
#     nga_7_subset$mean_may20_prec) / 12)

# Re-examine CHIRPS Rainfall data -------------------------------------------------------------
# Result : Re-extracted rainfall data after eliminating negative values. Coding provided in 
# CHIRPS_Nigeria_data_manipulation.R script modified on 8-3-20. 
# File is found here:
# /Volumes/Samsung T5/LACIE MacOS Extended/Tulane Research Projects/Malaria Consortium/Methodology/R Code/MC R Code/Malaria-Consortium/Malaria-Consortium
# Redo mean annual rainfall variable 8-3-20 since negative values were not removed previously
# or not converted to missing values
# TASK: revisit this code on 8-24-20
nga_7_subset$mean_an_rf <- ((nga_lga.df$mean_jun19_rf + nga_lga.df$mean_jul19_rf +
                                 nga_lga.df$mean_aug19_rf + nga_lga.df$mean_sep19_rf + nga_lga.df$mean_oct19_rf +
                                 nga_lga.df$mean_nov19_rf + nga_lga.df$mean_dec19_rf + nga_lga.df$mean_jan20_rf +
                                 nga_lga.df$mean_feb20_rf + nga_lga.df$mean_mar20_rf + nga_lga.df$mean_apr20_rf +
                                 nga_lga.df$mean_may20_rf) / 12)

# Re-examine CHIRPS Rainfall data - ensure no negative values in mean annual rf calculations
# Good to go this time 
summary(nga_7_subset$mean_an_rf)
# Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
# 30.85   81.59  117.90  130.48  166.15  300.44 

# Calculate jenks for rainfall ------------------------------------------------
# obtain 5 classes of breaks for mean annual rainfall
getJenksBreaks(nga_7_subset$mean_an_rf, 5, subset = NULL) 
# [1]  30.85152  92.29017 146.79149 206.53586 300.44352

# 8-3-20 Note: no change in rainfall class jenks/breaks after removing 
# negative rainfall values

# Create mean annual rainfall classes ----------------------------------------
nga_7_subset$rf_class[nga_7_subset$mean_an_rf >  206.53586] <- 4                                            # classify as rank 4
nga_7_subset$rf_class[nga_7_subset$mean_an_rf <= 206.53586   & nga_7_subset$mean_an_rf > 146.79149 ] <- 3   # classify as rank 3
nga_7_subset$rf_class[nga_7_subset$mean_an_rf <= 146.79149   & nga_7_subset$mean_an_rf > 92.29017  ] <-  2  # classify as rank 2
nga_7_subset$rf_class[nga_7_subset$mean_an_rf <  92.29017]  <- 1                                            # classify as rank 1

# Created rainfall weight
nga_7_subset$rf_wt <- 0.78

# Calculate rainfall weighted score
nga_7_subset$rf_score <- nga_7_subset$rf_class * nga_7_subset$rf_wt  

# Calculate rainfall unweighted score
nga_7_subset$rf_eqwt_score <- nga_7_subset$rf_class * nga_7_subset$eq_wt 

# Subset made on 8-26-20
# Subset nga_7_subset to make revised rainfall map 
rf_vars <- c(1, 3, 70:75)
rf_vars_slim <- nga_7_subset[rf_vars]
View(rf_vars_slim)

# see rf scores
table(rf_vars_slim$rf_score, useNA = "always")
# 0.78 1.56 2.34 3.12 <NA> 
#  245  296  110  124    0 

# 8-26-20 try this merged dataset for new rainfall map
# Success! Worked appropriately on 8-26-20
nga_7_rf.spdf <- merge(adm2.nga, rf_vars_slim, by="GID_2", duplicateGeoms = T)

# Subset nga_7_subset even further in order to merge with master dataset ------
# used in map production
# Keep the following variables: GID_2, mean_an_rf, rf_class, rf_wt, 
# rf_score, rf_eqwt_score

# reverse smc class
# if(nga_7_subset$smc_score_weighted == 0){ nga_7_subset$smc_score_weighted_rev == 1 }
# if (nga_7_subset$smc_score_weighted == 0) nga_7_subset$smc_score_reverse == 1 else nga_7_subset$smc_score_reverse == 0

#!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!#
# TASK: modify this code and perform this task in R  #
#!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!#

# write.csv in order to manually reverse 0 and 1 values
write.csv(nga_7_subset,"D:\\LACIE MacOS Extended\\Tulane Research Projects\\Malaria Consortium\\Nigeria Data\\nga_7_subset.csv", row.names = TRUE)
# Bring csv back in after modifying SMC values/scores (reversing)
nga_7_subset_smc_rev <- read.csv("D:\\LACIE MacOS Extended\\Tulane Research Projects\\Malaria Consortium\\Nigeria Data\\nga_7_subset_smc_rev.csv", header = TRUE)

```

## Data management

```{r}
#******************************************************************************
# Merge temp suitability index data with nga_7_subset_smc_rev
#******************************************************************************
nga_7_subset_smc_rev_2 <- merge(nga_7_subset_smc_rev, temp_suit.df, 
                    by="GID_2", duplicateGeoms = T)
# Make nga_7_subset_smc_rev_2 a data frame
# nga_7_subset_smc_rev_2.df <- as.data.frame(nga_7_subset_smc_rev_2)

```

## Temperature suitability jenks

```{r}
# --------------------------------------------------------------------------
# TEMP SUITABILITY jenks
# --------------------------------------------------------------------------
# Create jenks / breaks # temp suitability  ----------------------------------
# Obtain 5 classes of breaks for temp suitability
# For temp suitability, value near zero will represent low risk malaria, likely,
# The normalized Z(T) index of temperature suitability that incorporates not
# just the duration but also the degree of suitability across an average year. 
getJenksBreaks(nga_7_subset_smc_rev_2$mean_temp_suit, 5, subset = NULL)  
# [1] 0.3015740 0.4532296 0.5710385 0.6795321 0.7636224
```

## Temperature suitability classes

```{r}
# Create temp suitability classes
nga_7_subset_smc_rev_2$temp_suit_class[nga_7_subset_smc_rev_2$mean_temp_suit >  0.6795321] <- 4                                        # classify as rank 4
nga_7_subset_smc_rev_2$temp_suit_class[nga_7_subset_smc_rev_2$mean_temp_suit <= 0.6795321   & nga_7_subset_smc_rev_2$mean_temp_suit > 0.5710385 ] <- 3     # classify as rank 3
nga_7_subset_smc_rev_2$temp_suit_class[nga_7_subset_smc_rev_2$mean_temp_suit <= 0.5710385   & nga_7_subset_smc_rev_2$mean_temp_suit > 0.4532296 ] <- 2    # classify as rank 2
nga_7_subset_smc_rev_2$temp_suit_class[nga_7_subset_smc_rev_2$mean_temp_suit <  0.4532296] <- 1                                          # classify as rank 1

# Create slim subset to inspect temp suit class variables
temp_vars <- c(1, 3, 78, 79)
temp_vars_slim <- nga_7_subset_smc_rev_2[temp_vars]
View(temp_vars_slim)
```

## Temperature suitability score calculations

```{r}

# Create temp suitability weight ----------
nga_7_subset_smc_rev_2$temp_suit_wt <- 0.58 

# temp suit score calculation
nga_7_subset_smc_rev_2$temp_suit_score <- nga_7_subset_smc_rev_2$temp_suit_class * nga_7_subset_smc_rev_2$temp_suit_wt # create temp suitability weighted score 
# make temp suit sore a factor variable
nga_7_subset_smc_rev_2$temp_suit_score <- as.factor(nga_7_subset_smc_rev_2$temp_suit_score)
# Equal wt temp suitability score calculation
nga_7_subset_smc_rev_2$temp_suit_eq_wt_score <- nga_7_subset_smc_rev_2$temp_suit_class * nga_7_subset_smc_rev_2$eq_wt   #create temp suitability eq wt score

# NOTE: Disruptive event, which captures internally displaced populations 
# has a weight of 0.4 (already included in score (weighted)) 
# rather than original, 0.28

```

## PBO nets score calculations

```{r}

#******************************************************************************
#* TASK: Add PBO net designation layer to prioritization index and map
#* ----------------------------------------------------------------------------
#* PMI Nigeria will use results from insecticide resistance monitoring to
#* influence ITN procurement. Moderate to high levels of pyrethroid 
#* resistance have been identified over the last two years in two 
#* entomologic surveillance sites (Ebonyi and Oyo). PMI Nigeria plans to 
#* procure 1.7 million next-generation or PBO ITNs for Ebonyi State 
#* (2019 campaign). PMI Nigeria will work closely with PMI to design and 
#* implement an entomologic and epidemiologic evaluation of these nets. 
#* In Oyo State, PBO did not fully restore permethrin susceptibility, 
#* but did for deltamethrin. The next mass campaign in Oyo will occur in 2020
#* (Rescheduled for 2021 according to Mr. Okoko meeting 8/2020).
#* PMI Nigeria would like to procure next-generation or deltamethrin-based 
#* PBO ITNs for Oyo; however, with a need of five million ITNs next 
#* generation nets would require a subsidy or additional resources.
#* ---------------------------------------------------------------------------
#* From: PRESIDENT’S MALARIA INITIATIVE NIGERIA 
#* Malaria Operational Plan FY 2019
#* ---------------------------------------------------------------------------
#* "Acting upon the insecticide resistance data obtained from Ebonyi State, 
#* PMI plans to deploy PBO ITNs in 2019."
#* ---------------------------------------------------------------------------
#* Plans and justification for proposed activities with FY 2019 funding:
#* ----------------------------------------------------------------------------
#* "PMI Nigeria will use results from insecticide resistance monitoring to 
#* influence ITN procurement. Moderate to high levels of pyrethroid resistance 
#* have been identified over the last two years in two entomologic surveillance 
#* sites (Ebonyi and Oyo). PMI Nigeria plans to procure 1.7 million 
#* next-generation or PBO ITNs for Ebonyi State (2019 campaign)."
#* ****************************************************************************

# Create & set PBO class variable = 0 if PBO nets distributed in 2019 and
# 1 if PBO nets were not distributed in 2019, indicating that those regions
# are at higher risk for malaria, therefore, are assigned a higher value
nga_7_subset_smc_rev_2$PBO_class[nga_7_subset_smc_rev_2$NAME_1 == "Ebonyi"] <- 0
nga_7_subset_smc_rev_2$PBO_class[nga_7_subset_smc_rev_2$NAME_1 != "Ebonyi"] <- 1

# Create PBO weight
nga_7_subset_smc_rev_2$pbo_wt <- 0.53

# PBO score calculation
nga_7_subset_smc_rev_2$pbo_score <- nga_7_subset_smc_rev_2$PBO_class * nga_7_subset_smc_rev_2$pbo_wt # create pbo weighted score 

# Equal wt PBO score calculation
nga_7_subset_smc_rev_2$pbo_eq_wt_score <- nga_7_subset_smc_rev_2$PBO_class * nga_7_subset_smc_rev_2$eq_wt   #create pbo eq wt score

# Create slim PBO dataset to verify correct variable creation 
# and value assignment
# Create slim subset to inspect temp suit class variables
pbo_vars <- c(1, 3, 4, 84:87)
pbo_vars_slim <- nga_7_subset_smc_rev_2[pbo_vars]
View(pbo_vars_slim)
# n = 13 LGAs have value of 1, which is appropriate
```

## SMC class data manipulation

```{r}


# Create SMC class switch variable (reverse the values so that 1 implies greater -------
# malaria risk, and 0 implies less malaria risk)
nga_7_subset_smc_rev_2$smc_class_switch [nga_7_subset_smc_rev_2$SMC_class == 0  ] <- 1
nga_7_subset_smc_rev_2$smc_class_switch [nga_7_subset_smc_rev_2$SMC_class == 1  ] <- 0

# Added this bit of code after talking to Mr. Okoko - confirming SMC states
# Change SMC states (change for nga_7 and weighted spdf for final prioritization index)
# States that received SMC include:
# Borno         6% coverage
# Yobe         27% coverage
# Katsina      26% coverage
# Jigawa      107% coverage
# Sokoto      111% coverage
# Zamfara     114% coverage

nga_7_subset_smc_rev_2$smc_class_switch [nga_7_subset_smc_rev_2$NAME_1 == "Borno"   ] <- 0
nga_7_subset_smc_rev_2$smc_class_switch [nga_7_subset_smc_rev_2$NAME_1 == "Yobe"    ] <- 0
nga_7_subset_smc_rev_2$smc_class_switch [nga_7_subset_smc_rev_2$NAME_1 == "Katsina" ] <- 0
nga_7_subset_smc_rev_2$smc_class_switch [nga_7_subset_smc_rev_2$NAME_1 == "Jigawa"  ] <- 0
nga_7_subset_smc_rev_2$smc_class_switch [nga_7_subset_smc_rev_2$NAME_1 == "Sokoto"  ] <- 0
nga_7_subset_smc_rev_2$smc_class_switch [nga_7_subset_smc_rev_2$NAME_1 == "Zamfara" ] <- 0
```

## SMC score (re)calculations

```{r}

#!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!#
# TASK: Consider revising this section or
# combining it with other SMC section
#!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!#

# Recalculate smc_score_weighted -----------
nga_7_subset_smc_rev_2$smc_score_weighted <- nga_7_subset_smc_rev_2$smc_weight_0_82 * nga_7_subset_smc_rev_2$smc_class_switch
# Check that smc weighted score was calculated correctly - Looks good!
table(nga_7_subset_smc_rev_2$smc_score_weighted, useNA = "always")
#  0 0.82 <NA> # consider removing this 
# 72  703    0 
# After adding remaining states as SMC states the following
# results are created
#   0 0.82 <NA> 
# 143  632    0

#View slimmed nga_7_subset_smc_rev_2 for verification
smc_vars_check <- c(1, 3, 4, 49, 50, 94)
smc_vars_slim_check <- nga_7_subset_smc_rev_2[smc_vars_check]
View(smc_vars_slim_check)

# Merge smc_vars_slim_check with adm2
# use this for newest smc map 8-26-20
nga_7_smc_new.spdf <- merge(adm2.nga, smc_vars_slim_check, by="GID_2", duplicateGeoms = T)

# 8-3-20 convert temp_suit_score back to numeric for re-calculation
# Will need to run this prior to making final maps
nga_7_subset_smc_rev_2$temp_suit_score <- as.numeric(nga_7_subset_smc_rev_2$temp_suit_score)

```

## Data management steps

```{r}
# Bring in new rainfall and covid scores to malaria risk prioritization index
# convert spdf to data frames
nga_7_new.df <- as.data.frame(nga_7_new.spdf)
nga_7_jul21.df <- as.data.frame(nga_7_jul21.spdf)

# Slim datasets
rf_slim_vars <- c(1, 85)
nga_rf_slim <- nga_7_new.df[rf_slim_vars]
View(nga_rf_slim)

covid_slim_vars <- c(1, 136)
nga_covid_slim <- nga_7_jul21.df[covid_slim_vars]
View(nga_covid_slim)

# Convert scores back to numeric for weighted score sum
nga_rf_slim$rf_score <- as.numeric(nga_rf_slim$rf_score)
nga_covid_slim$COVID_case_score_jul_21_wtd <- as.numeric(nga_covid_slim$COVID_case_score_jul_21_wtd)
nga_7_subset_smc_rev_2$Built_score <- as.numeric(nga_7_subset_smc_rev_2$Built_score)
nga_7_subset_smc_rev_2$temp_suit_score <- as.numeric(nga_7_subset_smc_rev_2$temp_suit_score)

# rf_score from nga_7_new.spdf
# COVID_case_score_jul_21_wtd from nga_7_jul21.spdf

# Merge data sets
# Calculate scores

# Drop previous variables in nga_7_subset_smc_4_rev_2 that are no longer needed since
# they have been recalculated 8-27-20
nga_7_subset_smc_4_rev_3 <- nga_7_subset_smc_rev_2[ -c(21:23, 37:48, 50, 72:77, 88, 93:95, 97) ]

# slim rf and smc dataset then
rf_newest_slim_vars <- c(1, 3:7)
rf_newest <- rf_vars_slim[rf_newest_slim_vars]
View(rf_newest)

newest_smc_vars <- c(1, 16:18)
smc_newest <- nga_7_smc_new.df[newest_smc_vars]
View(smc_newest)

# Merge new rainfall(rf_vars_slim) and SMC dataset(nga_7_smc_new) with nga_7_subset_smc_4_rev_3 on 8-27-20
rf_smc_newest <- merge(x=rf_newest, y=smc_newest, by="GID_2", all = TRUE)
View(rf_smc_newest)

# Merge rfc_smc_newest with nga_7_subset_smc_4_rev_3
nga_7_subset_smc_4_rev_4 <- merge(x=rf_smc_newest, y=nga_7_subset_smc_4_rev_3, by="GID_2", all = TRUE)
View(nga_7_subset_smc_4_rev_4)

# Merge nga_covid_slim with nga_7_subset_smc_4_rev_4
nga_7_subset_smc_4_rev_5 <- merge(x=nga_covid_slim, y=nga_7_subset_smc_4_rev_4, by="GID_2", all = TRUE)
View(nga_7_subset_smc_4_rev_5)

# Modify COVID-19 case count variable and -------------------------------------
# years since last LLIN distribution variable then redo weighted NGA score
# Remove variables that are no longer necessary (previous COVID-19 case counts)
# Remove NAME_2.x variable

# slim nga_7_subset_smc_4_rev_5 data set to relevant variables
# remove irrelevant variables
nga_7_subset_smc_4_rev_6 <- nga_7_subset_smc_4_rev_5[-c(2, 11, 15:19, 25:27, 30:40, 45, 76)]

# Step: merge covid-19 case count relevant variables with ---------
# nga_7_subset_smc_4_rev_6 data set

# Slim covid dataset (updated 8-28-20)
covid_8_28_20_slim.df <- covid_8_28_20.df[-c(1,2,6,7)]

# Re-order columns
covid_8_28_20_slim.df <- covid_8_28_20_slim.df[,c(6,1,4,2,3,5)]

# Merge nga_7_subset_smc_4_rev_6 with covid_8_28_20_slim.df
nga_7_subset_smc_4_rev_7 <- merge(x=nga_7_subset_smc_4_rev_6, y=covid_8_28_20_slim.df, by="GID_2", all = TRUE)
names(nga_7_subset_smc_4_rev_7)
View(nga_7_subset_smc_4_rev_7)

# slim down nga_7_subset_smc_4_rev_7 data set to examine 
# number of years since last LLIN distribution variable
yrs_last_LLIN_8_28_20 <- nga_7_subset_smc_4_rev_7[c(1,11,15)]

# Re-examine nga_7.spdf

# convert nga_7.spdf to df
nga_7_new.df <- as.data.frame(nga_7.spdf)
# slim down nga_7_new.df to examine years since last LLIN distrib. variables
last_llin_8_28_20 <- nga_7_new.df[c(1, 25, 34:36)]
```

## ITN distribution

```{r}
# re-categorize the following states that were initi sched for 2020 distrib:
# Benue, Zamfara, Oyo, Osun, Adamawa, Kwara
last_llin_8_28_20$ITN_dist_class [last_llin_8_28_20$NAME_1.y == "Benue"   ] <- 2
last_llin_8_28_20$ITN_dist_class [last_llin_8_28_20$NAME_1.y == "Zamfara" ] <- 2
last_llin_8_28_20$ITN_dist_class [last_llin_8_28_20$NAME_1.y == "Oyo"     ] <- 2
last_llin_8_28_20$ITN_dist_class [last_llin_8_28_20$NAME_1.y == "Osun"    ] <- 2
last_llin_8_28_20$ITN_dist_class [last_llin_8_28_20$NAME_1.y == "Adamawa" ] <- 2
last_llin_8_28_20$ITN_dist_class [last_llin_8_28_20$NAME_1.y == "Kwara"   ] <- 2

# Create last ITN distribution weight variable
last_llin_8_28_20$last_itn_distrib_wt <- 0.75

# Update ITN_dist_score
last_llin_8_28_20$ITN_dist_score <- last_llin_8_28_20$ITN_dist_class * last_llin_8_28_20$last_itn_distrib_wt

# Create ITN_dist_score_factor variable
last_llin_8_28_20$ITN_dist_score_factor <- as.factor(last_llin_8_28_20$ITN_dist_score)

# Slim last_llin_8_28_20 data set to relevant variables prior to merge with
# nga_7_subset_smc_4_rev_7
last_llin_8_28_20_slim <- last_llin_8_28_20[c(1,3,4,5,6,7)]

# Remove previous last LLIN distribution related variables
nga_7_subset_smc_4_rev_8 <- nga_7_subset_smc_4_rev_7[-c(14:17)]

# Merge nga_7_subset_smc_4_rev_8 with last_llin_8_28_20_slim
nga_7_subset_smc_4_rev_9 <- merge(x=nga_7_subset_smc_4_rev_8, y=last_llin_8_28_20_slim, by="GID_2", all = TRUE)
names(nga_7_subset_smc_4_rev_9)
View(nga_7_subset_smc_4_rev_9)
```

## COVID score 

```{r}

# Create covid_case_score_wtd_numeric 
nga_7_subset_smc_4_rev_9$covid_case_score_wtd_numeric <- as.numeric(nga_7_subset_smc_4_rev_9$covid_case_score_wtd)

```

## Create final weighted scores

```{r}

#----------------------------------------------------------------------------------
# create final (weighted) NGA score -----------------------------------------------
#----------------------------------------------------------------------------------
nga_7_subset_smc_4_rev_9$weighted_score_sum <- nga_7_subset_smc_4_rev_9$rf_score + # Note This comes from dataset rf_vars_slim
    # now that rainfall data was re-calculated/redone after removing neg numbers on 8-27-20
    nga_7_subset_smc_4_rev_9$Built_score +        # OK
    nga_7_subset_smc_4_rev_9$smc_score_weighted + # OK
    nga_7_subset_smc_4_rev_9$covid_case_score_wtd_numeric + # updated on 8-28-20 
    nga_7_subset_smc_4_rev_9$weighted_net_avlblty_complement_score + # OK
    nga_7_subset_smc_4_rev_9$Disruptive_event_score + # OK
    nga_7_subset_smc_4_rev_9$ITN_dist_score +     # OK
    nga_7_subset_smc_4_rev_9$temp_suit_score +    # OK
    nga_7_subset_smc_4_rev_9$pbo_score            # OK

# Create factor variable for score sum for mapping
nga_7_subset_smc_4_rev_9$weighted_score_sum_factor <- as.factor(nga_7_subset_smc_4_rev_9$weighted_score_sum)

# Visualize variables in nga_7_subset_smc_4_rev_9
names(nga_7_subset_smc_4_rev_9)
```

## Create equally weighted scores

```{r}
#-------------------------------------------------------------------------
# Create equally weighed scores that weren't created previously-----------
# Note: Modified on 8-29-20
#-------------------------------------------------------------------------
# 
# # 1) smc_score_eq_weighted - ok
# nga_7_subset_smc_4_rev_9$smc_score_eq_weighted <- nga_7_subset_smc_4_rev_9$eq_wt * nga_7_subset_smc_4_rev_5$smc_class_switch
# 2) net availability eq wt score - ok
nga_7_subset_smc_4_rev_9$eq_wt_net_avlblty_complement_score <- nga_7_subset_smc_4_rev_9$eq_wt * nga_7_subset_smc_4_rev_9$net_avlblty_complement_class
# 3) COVID-19 equal wt score - ok
nga_7_subset_smc_4_rev_9$COVID_case_eq_wt_score <- nga_7_subset_smc_4_rev_9$eq_wt * nga_7_subset_smc_4_rev_9$COVID_class_aug_28
# 4) internally displaced pop equal wt score
nga_7_subset_smc_4_rev_9$idp_eq_wt_score      [nga_7_subset_smc_4_rev_9$Disruptive_event_score == 0.4] <- 0.5
nga_7_subset_smc_4_rev_9$idp_eq_wt_score      [nga_7_subset_smc_4_rev_9$Disruptive_event_score != 0.4] <- 0
# 5) itn distribution equal wt score
nga_7_subset_smc_4_rev_9$itn_dist_eq_wt_score [nga_7_subset_smc_4_rev_9$ITN_dist_score == 0.75  ] <- 0.5  # 1 x wt
nga_7_subset_smc_4_rev_9$itn_dist_eq_wt_score [nga_7_subset_smc_4_rev_9$ITN_dist_score == 1.5   ] <- 1   # 2 x wt
nga_7_subset_smc_4_rev_9$itn_dist_eq_wt_score [nga_7_subset_smc_4_rev_9$ITN_dist_score == 2.25  ] <- 1.5   # 3 x wt
nga_7_subset_smc_4_rev_9$itn_dist_eq_wt_score [nga_7_subset_smc_4_rev_9$ITN_dist_score == 3     ] <- 2     # 4 x wt
# 6) smc eq wt score (same calc as above but different variable name)
nga_7_subset_smc_4_rev_9$smc_eq_wt_score <- nga_7_subset_smc_4_rev_9$eq_wt * nga_7_subset_smc_4_rev_9$smc_class_switch
# 7) rf_eq_weighted score (using 8/29/20 corrected rainfall layer)
nga_7_subset_smc_4_rev_9$rf_eqwt_score <- nga_7_subset_smc_4_rev_9$eq_wt * nga_7_subset$rf_class
# 8) Equal wt PBO score calculation
nga_7_subset_smc_4_rev_9$pbo_eq_wt_score <- nga_7_subset_smc_4_rev_9$PBO_class * nga_7_subset_smc_4_rev_9$eq_wt     #create pbo eq wt score
# 9) Equal wt Built score calculation
nga_7_subset_smc_4_rev_9$Built_eq_wt_score <- nga_7_subset_smc_4_rev_9$Built_class * nga_7_subset_smc_4_rev_9$eq_wt #create built eq wt score
# 10) temp suitability
nga_7_subset_smc_4_rev_9$temp_suit_eq_wt_score <- nga_7_subset_smc_4_rev_9$temp_suit_class * nga_7_subset_smc_4_rev_9$eq_wt   #create temp suitability eq wt score

#---------------------------------------------------------------------------
# slim nga_7_subset_smc_rev_2 data set to examine equal wt score creation
#---------------------------------------------------------------------------
# Create slim subset to inspect temp suit class variables
# wt should be 0.75 for use in weighted score
#itn dist
itn_dist_vars <- c(1, 3, 4, 13, 14, 15, 70, 92 )
itn_dist_vars_slim <- nga_7_subset_smc_rev_2[itn_dist_vars]
View(itn_dist_vars_slim)
#smc
smc_vars <- c(1, 3, 4, 21, 22, 23, 49, 50, 88, 93, 94 )
smc_vars_slim <- nga_7_subset_smc_rev_2[smc_vars]
View(smc_vars_slim)
#rainfall
rf_vars <- c(1, 3, 72:76 )
rf_vars_slim <- nga_7_subset_smc_rev_2[rf_vars]
View(rf_vars_slim)
#net availability
net_avail_vars <- c(1, 3, 53, 61:64, 96 )
net_avail_vars_slim <- nga_7_subset_smc_rev_2[net_avail_vars]
View(net_avail_vars_slim)
#temp suitability
temp_vars <- c(1, 3, 78:82 )
temp_vars_slim <- nga_7_subset_smc_rev_2[temp_vars]
View(temp_vars_slim)
# built
built_vars <- c(1, 3, 55,65:71)
built_vars_slim <- nga_7_subset_smc_rev_2[built_vars]
View(built_vars_slim)
# pbo
pbo_vars <- c(1, 3, 55,83:87)
pbo_vars_slim <- nga_7_subset_smc_rev_2[pbo_vars]
View(pbo_vars_slim)
# COVID
covid_vars <- c(1, 3, 16, 17, 18, 51, 52, 90)
covid_vars_slim <- nga_7_subset_smc_rev_2[covid_vars]
View(covid_vars_slim)
# IDP
idp_vars <- c(1, 3, 19, 20, 91)
idp_vars_slim <- nga_7_subset_smc_rev_2[idp_vars]
View(idp_vars_slim)

# Dbl check indicator score calculations per TA's comments ----------------
# 1) SMC-       weighted (0.82) and equally weighted (0.5) - OK (0/1 binary)
# 2) RF -       weighted (0.78) and equally weighted (0.5) - OK
# 3) net avail- weighted (0.75) and equally weighted (0.5) - OK
# 4) ITN dist - weighted (0.75) and equally weighted (0.5) - OK
# 8) temp suit- weighted (0.58) and equally weighted (0.5) - OK
# 5) built -    weighted (0.54) and equally weighted (0.5) - OK
# 9) PBO -      weighted (0.53) and equally weighted (0.5) - OK (0/1 binary)
# 3) IDP -      weighted (0.40) and equally weighted (0.5) - OK (0/1 binary)
# 7) COVID -    weighted (0.28) and equally weighted (0.5) - OK

# convert factor to numeric
# nga_7_subset_smc_rev_2$temp_suit_eq_wt_score <- as.numeric(nga_7_subset_smc_rev_2$temp_suit_eq_wt_score)    # OK
    
```

## Create final unweighted scores

```{r}

# 8-4-20 need to recalculate these scores after sending weighted scores
# list to Alyssa
# create final (unweighted) NGA score --------------------------------------
nga_7_subset_smc_4_rev_9$eq_weighted_score_sum <- nga_7_subset_smc_4_rev_9$rf_eqwt_score + # OK
    nga_7_subset_smc_4_rev_9$Built_eq_wt_score +         # OK
    nga_7_subset_smc_4_rev_9$COVID_case_eq_wt_score +    # OK             
    nga_7_subset_smc_4_rev_9$eq_wt_net_avlblty_complement_score +  #OK
    nga_7_subset_smc_4_rev_9$idp_eq_wt_score + # OK # note renamed this "idp" vs disrupt.event
    nga_7_subset_smc_4_rev_9$itn_dist_eq_wt_score +      # OK                   
    nga_7_subset_smc_4_rev_9$smc_eq_wt_score +           # OK              
    nga_7_subset_smc_4_rev_9$temp_suit_eq_wt_score +     # OK
    nga_7_subset_smc_4_rev_9$pbo_eq_wt_score             # OK
    
# range of weighted score sum
summary(nga_7_subset_smc_4_rev_9$weighted_score_sum)
# Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
# 5.29    9.01   10.30   10.50   12.23   15.79 
# New range of weighted score sum as of 8-27-20
# Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
# 6.97   10.37   12.07   12.49   14.69   19.88 
#----------------------------------------------------
# Newest range of weighted score sum as of 8-27-20
# Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
# 6.43   10.54   12.82   12.63   14.82   19.85 
#----------------------------------------------------

# range of equally weighted score sum
summary(nga_7_subset_smc_4_rev_9$eq_weighted_score_sum)
# Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
# 5.500   8.500   9.500   9.665  11.000  14.500 
# revised equally weighted summary as of 8-5-20
# Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
# 7.19   11.51   12.76   13.05   14.76   20.76 
#----------------------------------------------------
# revised equally weighed summary on 8-29-20
# Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
# 4.500   7.500   8.000   8.335   9.500  12.000 
#----------------------------------------------------

```

## Jenks / natural breaks for final weighted NGA score sum

```{r}

# jenks for final weighted NGA score sum (updated 8-27-20) ---------------------------
getJenksBreaks(nga_7_subset_smc_4_rev_9$weighted_score_sum, 7, subset = NULL)  
# [1]  6.43  9.29 11.50 13.34 15.07 17.07 19.85

# rename final risk prioritization data set
prioritization.df <- nga_7_subset_smc_4_rev_9

```

## Create malaria (weighted) risk categories

```{r}

# Create nga malaria risk category (for weighted table)
prioritization.df$Malaria_risk[prioritization.df$weighted_score_sum >= 17.07 ] <- 6
prioritization.df$Malaria_risk[prioritization.df$weighted_score_sum >= 15.07 & prioritization.df$weighted_score_sum < 17.07 ] <- 5    # 
prioritization.df$Malaria_risk[prioritization.df$weighted_score_sum >= 13.34 & prioritization.df$weighted_score_sum < 15.07 ] <- 4    # 
prioritization.df$Malaria_risk[prioritization.df$weighted_score_sum >= 11.50 & prioritization.df$weighted_score_sum < 13.34 ] <- 3    # 
prioritization.df$Malaria_risk[prioritization.df$weighted_score_sum >= 9.29  & prioritization.df$weighted_score_sum < 11.50 ] <- 2    # 
prioritization.df$Malaria_risk[prioritization.df$weighted_score_sum >= 6.43  & prioritization.df$weighted_score_sum < 9.29  ] <- 1    # 

# Make malaria risk category a factor variable for chloropleth maps
prioritization.df$Malaria_risk_factor <- as.factor(prioritization.df$Malaria_risk)

# Verify no missing values for malaria risk factor variable -----
table(prioritization.df$Malaria_risk_factor, useNA = "always")
# 1    2    3    4    5    6 <NA> 
# 122  154  185  156  108   50    0 

# # distribution of equally weighted score sums (modified 8-5-20) ----------
# table(nga_7_subset_smc_rev_9$eq_weighted_score_sum, useNA = "always")
# # 5.5    6  6.5    7  7.5    8  8.5    9  9.5   10 10.5   11 11.5   12 12.5   13 13.5   14 14.5 
# #   1   18   13   36   40   60   77  116   80   80   44   37   51   39   42   20    7    8    6 

# Distribution of equally weighted score sum (modified 8-29-20)----------------------
table(prioritization.df$eq_weighted_score_sum, useNA = "always")
# 4.5    5  5.5    6  6.5    7  7.5    8  8.5    9  9.5   10 10.5   11 11.5   12 <NA> 
#   1   15   15   14   43   68  108  135  100   80   66   43   46   28   11    2    0 

```

## Jenks / natural breaks for equally weighted score sum

```{r}

# jenks for final equally weighted NGA score sum ---------------------------
getJenksBreaks(prioritization.df$eq_weighted_score_sum, 7, subset = NULL)  
# [1]  5.5  7.0  8.5  9.5 10.5 12.0 14.5 (previous jenks)
# [1]  4.5  6.0  7.0  8.0  9.0 10.0 12.0 (modified jenks on 8-29-20)

# Create nga malaria risk category (for equally weighted table)
prioritization.df$Malaria_risk_eq_wt[prioritization.df$eq_weighted_score_sum >= 10 ] <- 6
prioritization.df$Malaria_risk_eq_wt[prioritization.df$eq_weighted_score_sum >= 9.0  & prioritization.df$eq_weighted_score_sum < 10  ] <- 5   
prioritization.df$Malaria_risk_eq_wt[prioritization.df$eq_weighted_score_sum >= 8.0  & prioritization.df$eq_weighted_score_sum < 9.0 ] <- 4  
prioritization.df$Malaria_risk_eq_wt[prioritization.df$eq_weighted_score_sum >= 7.0  & prioritization.df$eq_weighted_score_sum < 8.0 ] <- 3   
prioritization.df$Malaria_risk_eq_wt[prioritization.df$eq_weighted_score_sum >= 6.0  & prioritization.df$eq_weighted_score_sum < 7.0 ] <- 2     
prioritization.df$Malaria_risk_eq_wt[prioritization.df$eq_weighted_score_sum >= 4.5  & prioritization.df$eq_weighted_score_sum < 6.0 ] <- 1    
```

## Data management steps

```{r}
# spot check distribution of equally weighted ---------------
table(prioritization.df$Malaria_risk_eq_wt, useNA = "always")
# something is not right here
#  1    2    3    4    5    6 <NA> 
# 31   57  176  235  146  130    0 

# spot check distribution of weighted------------------------
table(prioritization.df$Malaria_risk, useNA = "always")
# something is not right here
#   1    2    3    4    5    6 <NA> 
# 122  154  185  156  108   50    0 

#------------------------------------------------------------------------------------
# NOTE: NEED TO CHANGE THESE BACK TO NUMERIC (or factor) 
# IF MAKE ANOTHER PRIORITIZATION MAP
#------------------------------------------------------------------------------------

# Make equally weighted malaria risk category a factor variable for chloropleth maps
prioritization.df$Malaria_risk_eq_wt_factor <- as.factor(prioritization.df$Malaria_risk_eq_wt)
# Make built_score a factor variable
prioritization.df$Built_score <- as.factor(prioritization.df$Built_score)
# Make COVID_case_score_wtd a factor variable
prioritization.df$COVID_case_score_wtd <- as.factor(prioritization.df$COVID_case_score_wtd)
# Make weighted_net_avlblty_complement_score a factor variable
prioritization.df$weighted_net_avlblty_complement_score_factor <- as.factor(prioritization.df$weighted_net_avlblty_complement_score)
# Make pbo a factor variable
prioritization.df$pbo_score_factor <- as.factor(prioritization.df$pbo_score)

# Merge adm2.nga with nga_7 dataset by NAME_2 ---------------------------
# nga_master7 <- merge(adm2.nga, nga_master6, by="NAME_2", all = TRUE)

# Code added to include additional states receiving SMC after discussion with
# Mr. Okoko on 8-29-20

#---------------------------------------------------
# USE THIS OBJECT FOR MAPS 
#---------------------------------------------------
nga_7_8_29_20.spdf <- merge(adm2.nga, prioritization.df, by="GID_2", duplicateGeoms = T)

# Make nga_7.spdf a dataframe for closer inspection of n=775 elements
nga_7_8_29_20.df <- as.data.frame(nga_7_8_29_20.spdf)
# Note: GID_2 has 775 unique obs while NAME_2 has 770 unique obs. n =5 have duplicate names
# Bassa (2 entries), Irepodun (2 entries), Obi (2 entries), Ifelodun (2 entries),
# Surulere (2 entries)

# -----------------------------------------------------------------
# -----------------------------------------------------------------
View(nga_7_8_29_20.df) # use this data set as master as of 8-29-20
# -----------------------------------------------------------------
# -----------------------------------------------------------------

# Export weighted table for report -----------------
table_vars <- c(1, 5, 7, 77, 83 )
nga_7_8_29_20_slim.df <- nga_7_8_29_20.df[table_vars]
View(nga_7_8_29_20_slim.df)
# Create extra columns
nga_7_8_29_20_slim.df$Risk_Category <- nga_7_8_29_20_slim.df$Malaria_risk_factor
nga_7_8_29_20_slim.df$Prioritization_Level [nga_7_8_29_20_slim.df$Malaria_risk_factor == 6 ] <- "Extrememly high"
nga_7_8_29_20_slim.df$Prioritization_Level [nga_7_8_29_20_slim.df$Malaria_risk_factor == 5 ] <- "High"
nga_7_8_29_20_slim.df$Prioritization_Level [nga_7_8_29_20_slim.df$Malaria_risk_factor == 4 ] <- "Moderate high"
nga_7_8_29_20_slim.df$Prioritization_Level [nga_7_8_29_20_slim.df$Malaria_risk_factor == 3 ] <- "Moderate"
nga_7_8_29_20_slim.df$Prioritization_Level [nga_7_8_29_20_slim.df$Malaria_risk_factor == 2 ] <- "Moderate low"
nga_7_8_29_20_slim.df$Prioritization_Level [nga_7_8_29_20_slim.df$Malaria_risk_factor == 1 ] <- "Low"

# Change name of NAME_1.x to "State" ------------------------
nga_7_8_29_20_slim.df$State <- nga_7_8_29_20_slim.df$NAME_1.x
# Rename NAME_2.x variable to "LGA" -------------------------
nga_7_8_29_20_slim.df$LGA <- nga_7_8_29_20_slim.df$NAME_2.x
View(nga_7_8_29_20_slim.df)
# Drop unnecessary variables
nga_7_8_29_20_slim_2.df <- nga_7_8_29_20_slim.df[c(1,6,7,4,8,9)]
View(nga_7_8_29_20_slim_2.df)

# Convert factor to numeric for sorting -------------
nga_7_8_29_20_slim_2.df$Risk_Category <- as.numeric(nga_7_8_29_20_slim_2.df$Risk_Category)
# Sort by Risk category
nga_7_8_29_20_slim_2_sorted.df <- nga_7_8_29_20_slim_2.df[order(-nga_7_8_29_20_slim_2.df$Risk_Category),]    
View(nga_7_8_29_20_slim_2_sorted.df)

# Export table weighted scores, prioritization category, population
# for Alyssa on 8-29-20
write.csv(nga_7_8_29_20.df, "/Volumes/Samsung T5/LACIE MacOS Extended/Tulane Research Projects/Malaria Consortium/Nigeria Data/Datasets/nga_7_8_29_20.df.csv", row.names = TRUE)
write.csv(nga_7_8_29_20_slim_2_sorted.df, "/Volumes/Samsung T5/LACIE MacOS Extended/Tulane Research Projects/Malaria Consortium/Nigeria Data/Datasets/nga_7_8_29_20_slim_2_sorted.df.csv", row.names = TRUE) # exported on 8-29-20 @ 2050 hrs PST
# 
# # Rename variables
# nga_7_slim.df$Risk_category <- nga_7_slim.df$Malaria_risk_factor
# # drop Malaria_risk_factor
# table_vars_2 <- c(1, 2, 3, 4, 6)
# nga_7_slim_2.df <- nga_7_slim.df[table_vars_2]
# View(nga_7_slim_2.df)

# # Convert factor to numeric for sorting -------------
# nga_7_slim_2.df$Risk_category <- as.numeric(nga_7_slim_2.df$Risk_category)
# # Sort by Risk category
# nga_7_slim_2_sorted.df <- nga_7_slim_2.df[order(-nga_7_slim_2.df$Risk_category),]    
# View(nga_7_slim_2_sorted.df)

# # Rename Variables
# nga_7_slim_2_sorted.df$State <- nga_7_slim_2_sorted.df$NAME_1.x
# nga_7_slim_2_sorted.df$LGA <- nga_7_slim_2_sorted.df$NAME_2.x
# View(nga_7_slim_2_sorted.df)

# # Keep relevant variables
# table_vars_3 <- c(1, 4, 5, 6, 7)
# nga_7_slim_3_sorted.df <- nga_7_slim_2_sorted.df[table_vars_3]
# View(nga_7_slim_3_sorted.df)
# # Reorder variables
# table_vars_4 <- c(1,4,5,2,3)
# nga_7_slim_4_sorted.df <- nga_7_slim_3_sorted.df[table_vars_4]
# View(nga_7_slim_4_sorted.df)

#---------------------------------------------
# export weighted prioritization table to manipulate 
# in excel and paste into word document (exported on 8-29-20)
#---------------------------------------------
# write.csv(nga_7_slim_4_sorted.df,"/Volumes/Samsung T5/LACIE MacOS Extended/Tulane Research Projects/Malaria Consortium/Nigeria Data/Datasets/nga_7_slim_4_sorted_df_v2.csv", row.names = TRUE)
# write.csv(nga_7_8_29_20_slim.df, "/Volumes/Samsung T5/LACIE MacOS Extended/Tulane Research Projects/Malaria Consortium/Nigeria Data/Datasets/nga_7_8_29_20_slim_df.csv", row.names = TRUE)

# export nga_7.df to csv for Alyssa to use for budgeting 7-23-20
# Re-exported on 8-5-20 for Alyssa budget
# write.csv(nga_7.df, "/Volumes/Samsung T5/LACIE MacOS Extended/Tulane Research Projects/Malaria Consortium/Nigeria Data/Datasets/nga_7_df.csv", row.names = TRUE)

#---------------------------------------------
# Export equally weighted table for report ---
#---------------------------------------------
table_eq_wt_vars <- c(1, 5, 7, 81, 84)
eq_wt.df <- nga_7_8_29_20.df[table_eq_wt_vars]
View(eq_wt.df) # dataset numbers look good

# rename variables -------
eq_wt.df$Risk_category <- eq_wt.df$Malaria_risk_eq_wt
# drop Malaria_risk_factor
table_eq_wt_vars_2 <- c(1, 2, 3, 4, 6)
eq_wt_slim.df <- eq_wt.df[table_eq_wt_vars_2]
View(eq_wt_slim.df)

# spot check -------
table(eq_wt_slim.df$Risk_category, useNA = "always")
# Former distribution of risk categories
#  1    2    3    4    5    6 <NA> 
# 32  136  193  160  132  122    0 
# Distribution of risk categories as of 8-29-20
#  1    2    3    4    5    6 <NA> 
# 31   57  176  235  146  130    0 

# convert to numeric for sorting
# eq_wt_slim.df$Risk_category <- as.numeric(eq_wt_slim.df$Risk_category)

# Sort by risk category ------------
eq_wt_slim_sorted.df <- eq_wt_slim.df[order(-eq_wt_slim.df$Risk_category),]

```

## Assign prioritization levels to risk categroies (equally weighted)

```{r}

# Assign Prioritization level to risk category (equally weighted)
# eq_wt_slim_sorted.df$Risk_category[eq_wt_slim_sorted.df$Risk_category == "Extremely high"] <- 6
eq_wt_slim_sorted.df$Prioritization_level[eq_wt_slim_sorted.df$Risk_category == 6] <- "Extremely high"
eq_wt_slim_sorted.df$Prioritization_level[eq_wt_slim_sorted.df$Risk_category == 5] <- "High"
eq_wt_slim_sorted.df$Prioritization_level[eq_wt_slim_sorted.df$Risk_category == 4] <- "Moderate high"
eq_wt_slim_sorted.df$Prioritization_level[eq_wt_slim_sorted.df$Risk_category == 3] <- "Moderate"
eq_wt_slim_sorted.df$Prioritization_level[eq_wt_slim_sorted.df$Risk_category == 2] <- "Moderate low"
eq_wt_slim_sorted.df$Prioritization_level[eq_wt_slim_sorted.df$Risk_category == 1] <- "Low"

# Spot check
names(eq_wt_slim_sorted.df)
# [1] "GID_2"                "NAME_1.x"             "NAME_2.x"             "Risk_category"        "Prioritization_level"
head(eq_wt_slim_sorted.df)
# GID_2 NAME_1.x          NAME_2.x Risk_category Prioritization_level
# 1  NGA.1.1_1     Abia         Aba North             6       Extremely high
# 10 NGA.1.2_1     Abia         Aba South             6       Extremely high
# 11 NGA.1.3_1     Abia          Arochukw             6       Extremely high
# 12 NGA.1.4_1     Abia             Bende             6       Extremely high
# 13 NGA.1.5_1     Abia           Ikwuano             6       Extremely high
# 14 NGA.1.6_1     Abia Isiala Ngwa North             6       Extremely high
table(eq_wt_slim_sorted.df$Risk_category, useNA = "always")
# Previous risk category distribution
#  1    2    3    4    5    6 <NA>  # looks good
# 32  136  193  160  132  122    0 
# Risk category distribution as of 8-29-20
#  1    2    3    4    5    6 <NA> 
# 31   57  176  235  146  130    0 

table(eq_wt_slim_sorted.df$Prioritization_level, useNA = "always")
# Extremely high           High            Low       Moderate  Moderate high   Moderate low           <NA> 
#     130            146             31            176            235             57              0 

```

## Data management

```{r}
# Rename variable names
eq_wt_slim_sorted.df$State <- eq_wt_slim_sorted.df$NAME_1.x
eq_wt_slim_sorted.df$LGA <- eq_wt_slim_sorted.df$NAME_2.x
eq_wt_slim_sorted.df$Unweighted_score <- eq_wt_slim_sorted.df$eq_weighted_score_sum

head(eq_wt_slim_sorted.df)

# Reorder variables and remove duplicates
eq_weighted_vars_8_29_20 <- c(1,8,9,7,5,6)
eq_wt_slim_sort_2.df <- eq_wt_slim_sorted.df[eq_weighted_vars_8_29_20]

# spot check
table(eq_wt_slim_sort_2.df$Risk_category)
# 1   2   3   4   5   6 
# 31  57 176 235 146 130 

table(eq_wt_slim_sort_2.df$Prioritization_level)
# Extremely high           High            Low       Moderate  Moderate high   Moderate low 
#            130            146             31            176            235             57 

# View equally weighted dataset
View(eq_wt_slim_sort_2.df)

# Sort by Risk category
# nga_7_slim_2_sorted.df <- nga_7_slim_2.df[order(-nga_7_slim_2.df$Risk_category),]    
 # View(nga_7_slim_2_sorted.df)

# Export table to manipulate in excel and paste into word document
# Note: Exported on 8-29-20
write.csv(eq_wt_slim_sort_2.df,"/Volumes/Samsung T5/LACIE MacOS Extended/Tulane Research Projects/Malaria Consortium/Nigeria Data/Datasets/eq_wt_slim_sort_2_8_29_20.csv", row.names = TRUE)

# merge equally weighted dataset with -----------------------------------------
# adm2.nga to create spdf or convert to sf
eq_weighted_8_29_20.spdf <- merge(adm2.nga, eq_wt_slim_sort_2.df, by="GID_2", duplicateGeoms = T)

# convert risk category to factor (unweighted) ----------
eq_weighted_8_29_20.spdf$Risk_category <- as.factor(eq_weighted_8_29_20.spdf$Risk_category)

#----------------------------------------------------------------------------
# Make  spatial data frame of weighted data set -----------------------------
# merge weighted dataset with adm2.nga to verify mapping ----------
weighted.spdf <- merge(adm2.nga, nga_7_8_29_20_slim_2_sorted.df, by="GID_2", duplicateGeoms = T)

# convert risk category to factor (weighted) on 8-29-20 ----------
weighted.spdf$Risk_Category <- as.factor(weighted.spdf$Risk_Category)

# spot check
table(weighted.spdf$Risk_Category, useNA = "always")
# Previous distribution of weighted risk category
#   1    2    3    4    5    6 <NA> 
# 172  156  170   95  114   68    0 

# Distribution of weighted risk category as of 8-29-20
#   1    2    3    4    5    6 <NA> 
# 122  154  185  156  108   50    0 
```

## Create weighted prioritization / malaria risk map

###!Consider revising this title to "malaria intervention prioritization map"!
###! or something like that!

```{r}
#---------------------------------------------------------------------------
# --------------------------------------------------------------------------
# 1a) Create Weighted Priortization/Malaria Risk map -----------------------
# Status: complete on 7-21-20
# Note: Recreated on 8-4-20?
# Note: Modified on 8-29-20
# --------------------------------------------------------------------------
# create custom color palette
# magma <- viridisLite::magma(6)
# plasma <- viridisLite::plasma(6)
# inferno <- viridisLite::inferno(6)

# make map within plot window for check
# tm_shape(weighted.spdf) +
#     tm_fill("Risk_Category", title = "Nigeria Priority Areas", n = 6, style = "cat",
#             palette = "Blues", 
#             labels = c("Low", "Moderate Low", "Moderate", "Moderate High", "High", "Extremely High"))  + 
#     tm_polygons("GID_2", lwd = 0.25, col = "gray3", border.col = "grey50")  +
#     tm_scale_bar() + tm_compass(position = c("RIGHT","TOP")) +
#     tm_layout(frame = FALSE, legend.outside = TRUE, 
#               legend.position = c("RIGHT","BOTTOM"), legend.title.size = 1) +
#     tm_shape(adm1.nga) + tm_borders(col = "grey40", lwd = 1)



# Assign map above to object (grey50 border color and grey50 tm_borders)
nga_prioritization_map_blue_factor_legend_pos <- tm_shape(weighted.spdf) +
    tm_fill("Risk_Category", title = "Nigeria Priority Areas", n = 6,
            style = "cat", palette = "Blues",
            labels = c("Low", "Moderate Low", "Moderate",
                      "Moderate High", "High", "Extremely High"))  +
    tm_polygons("GID_2", lwd = 0.10, col = "gray3", border.col = "grey50")  +
    tm_scale_bar() + tm_compass(position = c("RIGHT","TOP")) +
    tm_layout(frame = FALSE, legend.outside = FALSE,
              legend.position = c(0.75,0.10), legend.title.size = 1) +
    tm_shape(adm1.nga) + tm_borders(col = "grey40", lwd = 0.35)

# Save weighted prioritization map
# setwd("D:\\LACIE MacOS Extended\\Tulane Research Projects\\Malaria Consortium\\Nigeria Data\\Maps\\prioritization weighted") # PC
setwd("/Volumes/Samsung T5/LACIE MacOS Extended/Tulane Research Projects/Malaria Consortium/Nigeria Data/Maps/prioritization weighted") # Mac
# tmap_save(nga_prioritization_map, "nga_prioritization_map.png",
#           width =6.78, height=5, units ='in', asp = 0)
tmap_save(nga_prioritization_map_blue_factor_legend_pos, "nga_prioritization_map_blue_factor_legend_pos_v5.png",
          width =6.78, height=5, units ='in', asp = 0)
```

### Create equally weighted prioritization / malaria risk map

```{r}


# --------------------------------------------------------------------------
# 1b) Create Equally weighted Prioritization/Malaria Risk map ---------------
# Status: complete on 7-21-20
# Note: recreated on 8-29-20
# --------------------------------------------------------------------------
# 
# # make map within plot window for check
# tm_shape(eq_weighted_8_29_20.spdf) +
#     tm_fill("Risk_category", title = "Nigeria Priority Areas", n = 6, style = "cat",
#             palette = "Purples", labels = c("Low", "Moderate Low", "Moderate", "Moderate High", "High", "Extremely High"))  + 
#     tm_polygons("GID_2", lwd = 0.25, col = "gray3", border.col = "grey50")  +
#     tm_scale_bar() + tm_compass(position = c("RIGHT","TOP")) +
#     tm_layout(frame = FALSE, legend.outside = TRUE, 
#               legend.position = c("RIGHT","BOTTOM"), legend.title.size = 1) +
#     tm_shape(adm1.nga) + tm_borders(col = "grey40", lwd = 1)

# Assign map above to object (grey50 border color and grey50 tm_borders)
nga_prioritization_eq_wt_map <- tm_shape(eq_weighted_8_29_20.spdf) +
    tm_fill("Risk_category", title = "Nigeria Priority Areas", n = 6, style = "cat",
            palette = "Purples",
            labels = c("Low", "Moderate Low", "Moderate",
                       "Moderate High", "High", "Extremely High"))  + 
    tm_polygons("GID_2", lwd = 0.30, col = "gray3", border.col = "grey50")  +
    tm_scale_bar() + tm_compass(position = c("RIGHT","TOP")) +
    tm_layout(frame = FALSE, legend.outside = FALSE,
              legend.position = c(0.75,0.10), legend.title.size = 1) +
    tm_shape(adm1.nga) + tm_borders(col = "grey40", lwd = 0.50)

# Save weighted prioritization map
# setwd("D:\\LACIE MacOS Extended\\Tulane Research Projects\\Malaria Consortium\\Nigeria Data\\Maps\\prioritization weighted") # PC
setwd("/Volumes/Samsung T5/LACIE MacOS Extended/Tulane Research Projects/Malaria Consortium/Nigeria Data/Maps/prioritization equally weighted") # Mac
tmap_save(nga_prioritization_eq_wt_map, "nga_prioritization_eq_wt_map_v3.png",
          width =6.78, height=5, units ='in', asp = 0)

#------------------------------------------------------------------------------
# 8-29-20 Create spdf for master dataset for input maps
master_8_29_20.spdf <- merge(adm2.nga, nga_7_8_29_20.df, 
                             by="GID_2", duplicateGeoms = T)
#------------------------------------------------------------------------------

```

## Create temperature suitability index map

```{r}
# --------------------------------------------------------------------------
# 1c) Create temperature suitabilty index map ------------------------------
# Status: complete on 7-21-20
# --------------------------------------------------------------------------

# create custom color palette
YlOrBr <- get_brewer_pal("YlOrBr", n = 4)
# PuRd <- get_brewer_pal("YlGnBu", n = 4, contrast = c(0, 0.42))

# make map within plot window for check
# tm_shape(nga_8.spdf) +
#     tm_fill("Built_score", title = "Built up area\npresence index", n = 4, style = "cat",
#             palette = PuRd, labels = c("Very high", "High", "Moderate", "Low"))  + 
#     tm_polygons("GID_2", lwd = 0.25, col = "gray3", border.col = "grey50")  +
#     tm_scale_bar() + 
#     tm_layout(frame = FALSE, legend.outside = TRUE, 
#               legend.position = c("RIGHT","BOTTOM"), legend.title.size = 1) +
#     tm_shape(adm1.nga) + tm_borders(col = "grey40", lwd = 1)

# Assign map above to object (grey50 border color and grey50 tm_borders)
temp_suit_map <- tm_shape(nga_7.spdf) +
    tm_fill("temp_suit_score", title = "P.f.\ntemperature\nsuitability\nindex", n = 4, style = "cat",
            palette = YlOrBr, labels = c("Low", "Moderate", "High", "Very high"))  + 
    tm_polygons("GID_2", lwd = 0.10, col = "gray3", border.col = "grey50")  +
    tm_scale_bar(position = c(0.65, 0.01)) + 
    tm_layout(frame = FALSE, legend.outside = TRUE, 
              legend.position = c("RIGHT","BOTTOM"), legend.title.size = 1) +
    tm_shape(adm1.nga) + tm_borders(col = "grey40", lwd = 0.35)

# try confirmation map for temp suit using master_8_29_20.spdf
temp_suit_map <- tm_shape(master_8_29_20.spdf) +
    tm_fill("temp_suit_score", title = "P.f.\ntemperature\nsuitability\nindex", n = 4, style = "cat",
            palette = YlOrBr, labels = c("Low", "Moderate", "High", "Very high"))  + 
    tm_polygons("GID_2", lwd = 0.10, col = "gray3", border.col = "grey50")  +
    tm_scale_bar(position = c(0.65, 0.01)) + 
    tm_layout(frame = FALSE, legend.outside = TRUE, 
              legend.position = c("RIGHT","BOTTOM"), legend.title.size = 1) +
    tm_shape(adm1.nga) + tm_borders(col = "grey40", lwd = 0.35)

# Save Built up presence map
# setwd("D:\\LACIE MacOS Extended\\Tulane Research Projects\\Malaria Consortium\\Nigeria Data\\Maps\\built") # PC
setwd("/Volumes/Samsung T5/LACIE MacOS Extended/Tulane Research Projects/Malaria Consortium/Nigeria Data/Maps/temp suitability")
tmap_save(temp_suit_map, "temp_suit_map_8_29_20.png",
          width =6.78, height=5, units ='in', asp = 0)

```

## Create built up presence map

```{r}

# --------------------------------------------------------------------------
# 1d) Create Built up presence map -----------------------------------------
# Status: complete on 7-21-20
# --------------------------------------------------------------------------

# create custom color palette
YlGnBu <- get_brewer_pal("YlGnBu", n = 4, contrast = c(0, 0.42))

# make map within plot window for check
# tm_shape(nga_7.spdf) +
#     tm_fill("Built_score", title = "Built up area\npresence index", n = 4, style = "cat",
#             palette = PuRd, labels = c("Very high", "High", "Moderate", "Low"))  + 
#     tm_polygons("GID_2", lwd = 0.25, col = "gray3", border.col = "grey50")  +
#     tm_scale_bar() + 
#     tm_layout(frame = FALSE, legend.outside = TRUE, 
#               legend.position = c("RIGHT","BOTTOM"), legend.title.size = 1) +
#     tm_shape(adm1.nga) + tm_borders(col = "grey40", lwd = 1)

# Assign map above to object (grey50 border color and grey50 tm_borders)
nga_built_wtd_map <- tm_shape(master_8_29_20.spdf) +
    tm_fill("Built_score", title = "Built up\npresence\nindex", n = 4, style = "cat",
            palette = YlGnBu, labels = c("Very high", "High", "Moderate", "Low"))  + 
    tm_polygons("GID_2", lwd = 0.10, col = "gray3", border.col = "grey50")  +
    tm_scale_bar(position = c(0.65, 0.01)) + 
    tm_layout(frame = FALSE, legend.outside = TRUE, 
              legend.position = c("RIGHT","BOTTOM"), legend.title.size = 1) +
    tm_shape(adm1.nga) + tm_borders(col = "grey40", lwd = 0.35)

# Save Built up presence map
# setwd("D:\\LACIE MacOS Extended\\Tulane Research Projects\\Malaria Consortium\\Nigeria Data\\Maps\\built") # PC
setwd("/Volumes/Samsung T5/LACIE MacOS Extended/Tulane Research Projects/Malaria Consortium/Nigeria Data/Maps/built") # Mac
tmap_save(nga_built_wtd_map, "nga_built_wtd_map_8_29_20.png",
          width =6.78, height=5, units ='in', asp = 0)

```

## Create COVID-19 map

```{r}

# --------------------------------------------------------------------------
# 2) Create COVID map ------------------------------------------------------
# Map modified on 8-29-20
# --------------------------------------------------------------------------

#create COVID spdf ------------------------------------------------------------
covid_8_28_20.spdf <- merge(adm2.nga, covid_8_28_20.df, by="GID_2", 
                            duplicateGeoms = T)

# Assign map to object * used in tmap_arrange: covid_case_score_wtd_map
# 5  1142  2555  5094 18056 # new case count categories for 
# Aug 8, 2020 NGA

# covid_8_28_20.df$COVID_class_aug_28[covid_8_28_20.df$Covid_case_28_8_20 > 5094] <- 4                                               # classify as rank (class) 4
# covid_8_28_20.df$COVID_class_aug_28[covid_8_28_20.df$Covid_case_28_8_20 <= 5094 & covid_8_28_20.df$Covid_case_28_8_20 > 2555] <- 3 # classify as rank (class) 3
# covid_8_28_20.df$COVID_class_aug_28[covid_8_28_20.df$Covid_case_28_8_20 <= 2555 & covid_8_28_20.df$Covid_case_28_8_20 > 1142] <- 2 # classify as rank (class) 2
# covid_8_28_20.df$COVID_class_aug_28[covid_8_28_20.df$Covid_case_28_8_20 <= 1142] <- 1  

covid_case_score_wtd_map <- 
    tm_shape(master_8_29_20.spdf) + 
    tm_fill("covid_case_score_wtd", title = "COVID-19\ncase counts\nas of\nAug 28, 2020", n = 4, style = "cat", 
            palette = "YlOrRd", labels = c("< 1143", "1,143 - 2,555", "2,556 - 5,094", "> 5,094"), 
            lwd =0.01, border.col = "black") +
    tm_polygons(lwd = 0.25, col = "gray3")  + # "GID_2", border.col = "grey50"
    tm_scale_bar(position = c(0.65, 0.01)) + 
    tm_layout(frame = FALSE, legend.outside = TRUE, 
              legend.position = c("RIGHT","BOTTOM"), legend.title.size = 1) +
    tm_shape(adm1.nga) + tm_borders(col = "grey40", lwd = 1)

# Save COVID map
# setwd("D:\\LACIE MacOS Extended\\Tulane Research Projects\\Malaria Consortium\\Nigeria Data\\Maps\\covid") # PC
setwd("/Volumes/Samsung T5/LACIE MacOS Extended/Tulane Research Projects/Malaria Consortium/Nigeria Data/Maps/covid")
tmap_save(covid_case_score_wtd_map, "covid_score_aug_28_wtd_map.png",
          width =6.78, height=5, units ='in', asp = 0)
# tmap_save(covid_wtd_map_2, "covi_wtd_map_2.png",
#           width =6.78, height=5, units ='in', asp = 0)
    
```

## Create mean annual rainfall map

```{r}
# -------------------------------------------------------------------------
# Mean annual rainfall map Nigeria modified 8-26-20 -----------------------
# -------------------------------------------------------------------------
    mean_annual_rf_map <- tm_shape(nga_7_rf.spdf) + 
        tm_fill("rf_score", title = "Mean annual\nrainfall (mm)",
                n = 4, style = "cat", 
                palette = "Greens", labels = c("\U2264 92 mm", "91 - 147 mm ", "148 - 207 mm", "> 207 mm"),
                lwd =0.01, border.col = "black") + 
        tm_polygons(lwd = 0.25, col = "gray3") +
        tm_scale_bar(position = c(0.65, 0.01)) + 
        tm_layout(frame = FALSE, legend.outside = TRUE, 
                  legend.position = c("RIGHT","BOTTOM"), legend.title.size = 1) +
        tm_shape(adm1.nga) + tm_borders(col = "grey40", lwd = 1)
    
# Save Rainfall map modified 8-26-20 -----------------------
# setwd("D:\\LACIE MacOS Extended\\Tulane Research Projects\\Malaria Consortium\\Nigeria Data\\Maps\\covid") # PC
setwd("/Volumes/Samsung T5/LACIE MacOS Extended/Tulane Research Projects/Malaria Consortium/Nigeria Data/Maps/rainfall")
tmap_save(mean_annual_rf_map, "nga_rainfall_map_2.png",
          width =6.78, height=5, units ='in', asp = 0)

```

## Create net availability map

```{r}

# --------------------------------------------------------------------------
# 3) Create net availability variable and map ---------------------------------
# --------------------------------------------------------------------------
# STATUS: GOOD TO GO AS OF 7-15-20
# [1] 0.009014783 0.233290972 0.354107688 0.493038245 0.792025051
# (to use in legend)
# [1] 0.2079749 0.4951487 0.6451532 0.7655778 0.9909852

# unicode character
# U+2664 is less-than or equal to
# U+2265 is greater-than or equal to

# tm_shape(nga_7.spdf) + 
#     tm_fill("net_avlblty_complement_class", title = "ITN Availability", n = 4, style = "cat", 
#             palette = "Purples", labels = c(">67%", "50-67%", "35-50%", "<35%"), 
#             lwd =0.01, border.col = "black") + 
#     tm_polygons(lwd = 0.25, col = "gray3") +
#     tm_scale_bar() + tm_compass( position = c("left","top")) +
#     tm_layout(frame = FALSE, legend.outside = TRUE, 
#               legend.outside.position = "right")

RdPu <- get_brewer_pal("RdPu", n = 4, contrast = c(0, 0.42))

# assign map to object
ITN_avlblty_wt_map <- tm_shape(nga_7.spdf) + 
    tm_fill("weighted_net_avlblty_complement_score_factor", title = "ITN Availability", n = 4, style = "cat", 
            palette = "RdPu", labels = c("\u2265 49%", "35 - 48%", "23 - 34%", "<23%"), 
            lwd =0.01, border.col = "black") + 
    tm_polygons("GID_2", lwd = 0.10, col = "gray3", border.col = "grey50")  +
    tm_scale_bar(position = c(0.65, 0.01)) + 
    tm_layout(frame = FALSE, legend.outside = TRUE, 
              legend.position = c("RIGHT","BOTTOM"), legend.title.size = 1) +
    tm_shape(adm1.nga) + tm_borders(col = "grey40", lwd = 0.35)

# Save itn availability map
# setwd("D:\\LACIE MacOS Extended\\Tulane Research Projects\\Malaria Consortium\\Nigeria Data\\Maps\\itn availability") # PC
setwd("/Volumes/Samsung T5/LACIE MacOS Extended/Tulane Research Projects/Malaria Consortium/Nigeria Data/Maps/itn availability") # mac
tmap_save(ITN_avlblty_wt_map, "ITN_avlblty_wt_map.png",
          width =6.78, height=5, units ='in', asp = 0)

```

## Create SMC map

```{r}

# --------------------------------------------------------------------------
# 4) Create SMC map -----------------------------------------------------------
# --------------------------------------------------------------------------
# STATUS: GOOD TO GO AS OF 7-15-20
# tm_shape(hdx_nga_admin_0) + tm_polygons("ADM0_EN") + tm_legend(show=FALSE) + 
# How many LGAs are receiving SMC? n = 72
# table(nga_master6$SMC_class)
#       0    1 <NA> 
#     703   72    0 

# spot check SMC map
ltred <- get_brewer_pal("-Reds", n = 3, contrast = c(0.04, 0.44))

# tm_shape(nga_7.spdf) + 
#     tm_fill("smc_score_weighted", title = "SMC LGAs ", n=2, style = "cat", 
#             palette = ltred, labels = c("No", "Yes"), 
#             lwd =0.01, border.col = "black") + 
#     tm_polygons(lwd = 0.5, col = "gray3") +
#     tm_scale_bar(position = c(0.65, 0.01)) + 
#     tm_layout(frame = FALSE, legend.outside = TRUE, 
#               legend.position = c("RIGHT","BOTTOM"), legend.title.size = 1,
#               aes.palette = list(seq = "-Reds")) + 
#     tm_shape(adm1.nga) + tm_borders(col = "grey40", lwd = 1) +

# assign SMC map to object
nga_smc_weighted_map <- tm_shape(nga_7.spdf) + 
    tm_fill("smc_score_weighted", title = "SMC LGAs ", n=2, style = "cat", 
            palette = ltred, labels = c("No", "Yes"), 
            lwd =0.01, border.col = "black") + 
    tm_polygons(lwd = 0.5, col = "gray3") +
    tm_scale_bar(position = c(0.65, 0.01)) + 
    tm_layout(frame = FALSE, legend.outside = TRUE, 
              legend.position = c("RIGHT","BOTTOM"), legend.title.size = 1,
              aes.palette = list(seq = "-Reds")) +
    tm_shape(adm1.nga) + tm_borders(col = "grey40", lwd = 1) +
    tm_add_legend(reverse = TRUE)

# convert nga_7_smc_new.spdf to df for quick check
nga_7_smc_new.df <- as.data.frame(nga_7_smc_new.spdf)
View(nga_7_smc_new.df)

# add character variable for map
# nga_7_smc_new.spdf$smc_score_weighted_fac <- as.factor(nga_7_smc_new.spdf$smc_score_weighted)

# assign newest (8-26-20) SMC map to object
nga_smc_weighted_map_new <- tm_shape(nga_7_smc_new.spdf) + 
    tm_fill("smc_class_switch", title = "SMC LGAs ", n=2, style = "cat", 
            palette = ltred, labels = c("Yes", "No"), 
            lwd =0.01, border.col = "black") + 
    tm_polygons(lwd = 0.5, col = "gray3") +
    tm_scale_bar(position = c(0.65, 0.01)) + 
    tm_layout(frame = FALSE, legend.outside = TRUE, 
              legend.position = c("RIGHT","BOTTOM"), legend.title.size = 1,
              aes.palette = list(seq = "-Reds")) +
    tm_shape(adm1.nga) + tm_borders(col = "grey40", lwd = 1) 

# # no legend outside?
# nga_smc_weighted_map_arrange <- tm_shape(nga_master7.spdf) + 
#     tm_fill("smc_score_weighted", title = "SMC LGAs ", n=2, style = "cat", 
#             palette = "Reds", labels = c("No", "Yes"), 
#             lwd =0.01, border.col = "black") + 
#     tm_polygons(lwd = 0.5, col = "gray3") +
#     #tm_scale_bar() + tm_compass( position = c("left","top")) +
#     tm_layout(frame = FALSE)#, legend.outside = TRUE, legend.outside.position = "right")

# try water layer
# nga_water <- tm_shape(adm0.nga) + tm_fill(col="lightblue")=

# Save SMC map
# setwd("D:\\LACIE MacOS Extended\\Tulane Research Projects\\Malaria Consortium\\Nigeria Data\\Maps\\smc") # PC
setwd("/Volumes/Samsung T5/LACIE MacOS Extended/Tulane Research Projects/Malaria Consortium/Nigeria Data/Maps/smc") # Mac
# tmap_save(nga_smc_weighted_map, "nga_smc_weighted_map_2.png",
#           width =6.78, height=5, units ='in', asp = 0)
tmap_save(nga_smc_weighted_map_new, "nga_smc_weighted_map_new.png",
          width =6.78, height=5, units ='in', asp = 0)

```

## Create disruptive event map

```{r}

# --------------------------------------------------------------------------
# 5) Create disruptive event map ----------------------------------------------
# Internally displaced population ------------------------------------------
# --------------------------------------------------------------------------
# tm_shape(nga_master7.spdf) +
#     tm_fill("Disruptive_event_score", title = "Disruptive event\nInternally displaced,\nHigh risk populations", n=2, style = "cat",
#             palette = "Oranges", labels = c("No", "Yes"),
#             lwd =0.01, border.col = "black") +
#     tm_polygons(lwd = 0.5, col = "gray3") +
#     tm_scale_bar() + 
#     tm_layout(frame = FALSE, legend.outside = TRUE,
#               legend.position = c("RIGHT","BOTTOM"), legend.title.size = 1) +
#     tm_shape(adm1.nga)  + tm_borders(col = "grey40", lwd = 1)

# assign map to object
disruptive_event_weighted_map <- tm_shape(nga_7.spdf) + 
    tm_fill("Disruptive_event_score", title = "Internally\ndisplaced\npopulations", n=2, style = "cat", 
            palette = "Oranges", labels = c("No", "Yes"), 
            lwd =0.01, border.col = "black") + 
    tm_polygons(lwd = 0.5, col = "gray3") +
    tm_scale_bar(position = c(0.65, 0.01)) + 
    tm_layout(frame = FALSE, legend.outside = TRUE, 
              legend.position = c("RIGHT","BOTTOM"), legend.title.size = 1) +
    tm_shape(adm1.nga) + tm_borders(col = "grey40", lwd = 1)

# # no scale bar or compass
# disruptive_event_weighted_map_arrange <- tm_shape(nga_master7.spdf) +
#     tm_fill("Disruptive_event_score", title = "Disruptive event\nInternally displaced,\nHigh risk populations", n=2, style = "cat", 
#             palette = "Oranges", labels = c("No", "Yes"), 
#             lwd =0.01, border.col = "black") + 
#     tm_polygons(lwd = 0.5, col = "gray3") +
#     tm_layout(frame = FALSE)

# Save disruptive event map
setwd("D:\\LACIE MacOS Extended\\Tulane Research Projects\\Malaria Consortium\\Nigeria Data\\Maps\\disruptive event") #pc
# setwd("/Volumes/Samsung T5/LACIE MacOS Extended/Tulane Research Projects/Malaria Consortium/Nigeria Data/Maps/disruptive event") #mac
tmap_save(disruptive_event_weighted_map, "disruptive_event_weighted_map.png",
          width =6.78, height=5, units ='in', asp = 0)

```

## Create Last LLIN campaign map

```{r}

# --------------------------------------------------------------------------
# 6) Create LLIN last LLIN campaign map ---------------------------------------
# --------------------------------------------------------------------------

# quick check for missing values
table(nga_master6_sorted$ITN_dist_score, useNA = "always")
#    0.75  1.5 2.25    3 <NA> 
#     512  107   17  139    0 

# tm_shape(nga_master7.spdf) + 
#     tm_fill("ITN_dist_score", title = "Years since last\nLLIN distribution",
#             n = 4, style = "cat", 
#             palette = "Blues", labels = c("< 3 years", "3 years",
#                                           "5 years", ">=6 years"), 
#             lwd =0.01, border.col = "black") + 
#     tm_polygons(lwd = 0.25, col = "gray3") +
#     tm_scale_bar() + tm_compass( position = c("left","top")) +
#     tm_layout(frame = FALSE, legend.outside = TRUE, 
#               legend.outside.position = "right")

# assign map to object
yr_last_llin_wt_map <- tm_shape(master_8_29_20.spdf) + 
    tm_fill("ITN_dist_score_factor", title = "Years since last\nITN distribution",
            n = 4, style = "cat", 
            palette = "Blues", labels = c("< 3 years", "3 years",
                                          "5 years", "\u2265 6 years"), 
            lwd =0.01, border.col = "black") + 
    tm_polygons(lwd = 0.25, col = "gray3") +
    tm_scale_bar(position = c(0.65, 0.01)) + 
    tm_layout(frame = FALSE, legend.outside = TRUE, 
              legend.position = c("RIGHT","BOTTOM"), legend.title.size = 1) +
    tm_shape(adm1.nga) + tm_borders(col = "grey40", lwd = 1)

# Save last LLIN campaign map
# setwd("D:\\LACIE MacOS Extended\\Tulane Research Projects\\Malaria Consortium\\Nigeria Data\\Maps\\yr last LLIN") # pc
setwd("/Volumes/Samsung T5/LACIE MacOS Extended/Tulane Research Projects/Malaria Consortium/Nigeria Data/Maps/yr last LLIN") #mac
tmap_save(yr_last_llin_wt_map, "yr_last_llin_wt_map_8_29_20.png", 
          width =6.78, height=5, units ='in', asp = 0)

```

## Remove this section?

```{r}

# --------------------------------------------------------------------------
# 7) Create mean annual rainfall map ---------------------------------------
# --------------------------------------------------------------------------
# 
# # Re-create dataset for make new rainfall map
# nga_7_new.spdf <- merge(adm2.nga, nga_7_subset, by="GID_2", duplicateGeoms = T)
# # Convert score to factor
# nga_7_new.spdf$rf_score <- as.factor(nga_7_new.spdf$rf_score)
# 
# # Rainfall jenks are:
# # [1]  30.85152  92.29017 146.79149 206.53586 300.44352
# 
# # TASK: FIX LABELS FOR RAIN!!!
# 
# # Spot check rainfall map
# tm_shape(nga_7_new.spdf) + 
#     tm_fill("rf_score", title = "Mean annual\nrainfall (mm)",
#             n = 4, style = "cat", 
#             palette = "Greens", labels = c("\U2264 92 mm", "> 92 mm & \U2264 147 mm ", "> 147 mm & \U2264 207 mm", "> 207 mm"), 
# lwd =0.01, border.col = "black") + 
#     tm_polygons(lwd = 0.25, col = "gray3") +
#     tm_scale_bar(position = c(0.65, 0.01)) + 
#     tm_layout(frame = FALSE, legend.outside = TRUE, 
#               legend.position = c("RIGHT","BOTTOM"), legend.title.size = 1) +
#     tm_shape(adm1.nga) + tm_borders(col = "grey40", lwd = 1)
# 
# # Assign rf map to object
# mean_annual_rf_map <- tm_shape(nga_7_new.spdf) + 
#     tm_fill("rf_score", title = "Mean annual\nrainfall (mm)",
#             n = 4, style = "cat", 
#             palette = "Greens", labels = c("\U2264 92 mm", "91-147 mm ", "148-207 mm", "> 207 mm"),
#             lwd =0.01, border.col = "black") + 
#     tm_polygons(lwd = 0.25, col = "gray3") +
#     tm_scale_bar(position = c(0.65, 0.01)) + 
#     tm_layout(frame = FALSE, legend.outside = TRUE, 
#               legend.position = c("RIGHT","BOTTOM"), legend.title.size = 1) +
#     tm_shape(adm1.nga) + tm_borders(col = "grey40", lwd = 1)
# 
# # Save rainfall map
# # setwd("D:\\LACIE MacOS Extended\\Tulane Research Projects\\Malaria Consortium\\Nigeria Data\\Maps\\rainfall") # pc
# setwd("/Volumes/Samsung T5/LACIE MacOS Extended/Tulane Research Projects/Malaria Consortium/Nigeria Data/Maps/rainfall") #mac
# tmap_save(mean_annual_rf_map, "mean_annual_rf_map_new_2.png", 
#           width =6.78, height=5, units ='in', asp = 0)

```

## Create PBO map

```{r}
# --------------------------------------------------------------------------
# 5) Create PBO map ----------------------------------------------
# --------------------------------------------------------------------------

YlGn <- get_brewer_pal("YlGn", n = 2, contrast = c(0, 0.42))

# assign map to object
pbo_wtd_map <- tm_shape(nga_7.spdf) + 
    tm_fill("pbo_score_factor", title = "PBO nets\ndistributed\nin 2019", n=2, style = "cat", 
            palette = "YlGn", labels = c("Yes", "No"), 
            lwd =0.01, border.col = "black") + 
    tm_polygons("GID_2", lwd = 0.10, col = "gray3", border.col = "grey50")  +
    tm_scale_bar(position = c(0.65, 0.01)) + 
    tm_layout(frame = FALSE, legend.outside = TRUE, 
              legend.position = c("RIGHT","BOTTOM"), legend.title.size = 1) +
    tm_shape(adm1.nga) + tm_borders(col = "grey40", lwd = 0.35)

# Save disruptive event map
# setwd("D:\\LACIE MacOS Extended\\Tulane Research Projects\\Malaria Consortium\\Nigeria Data\\Maps\\disruptive event") #pc
setwd("/Volumes/Samsung T5/LACIE MacOS Extended/Tulane Research Projects/Malaria Consortium/Nigeria Data/Maps/pbo") #mac
tmap_save(pbo_wtd_map, "pbo_wtd_map.png",
          width =6.78, height=5, units ='in', asp = 0)

```

## Create faceted maps

```{r}

# --------------------------------------------------------------------------
# Make tmap_arrange figure -------------------------------------------------
# --------------------------------------------------------------------------
# Use tmap_arrange
# 3 column map with compass and scale bar
# tmap_arrange_compass_scale <- tmap_arrange(nga_states_map_0_75_text, nga_smc_weighted_map, 
#              disruptive_event_weighted_map, yr_last_llin_wt_map,
#              ITN_access_wt_map, covid_case_score_wtd_map, ncol = 3)

# Assign tmap arrange to object
# Map has compass and scale bar, 2 columns of maps ----
tmap_arrange_1 <- tmap_arrange(nga_smc_weighted_map_new, 
                               disruptive_event_weighted_map, 
                               yr_last_llin_wt_map, 
                               ITN_avlblty_wt_map,
                               covid_case_score_wtd_map,
                               nga_built_wtd_map,
                               ncol = 2)

tmap_arrange_2 <- tmap_arrange(mean_annual_rf_map,
                               temp_suit_map,
                               pbo_wtd_map,
                               ncol = 2)

# Save tmap_arrange 2 column outputs
# setwd("D:\\LACIE MacOS Extended\\Tulane Research Projects\\Malaria Consortium\\Nigeria Data\\Maps\\tmap_arrange") # PC
setwd("/Volumes/Samsung T5/LACIE MacOS Extended/Tulane Research Projects/Malaria Consortium/Nigeria Data/Maps/tmap_arrange") # Mac

# save 2 columns of tmap_arrange (6 maps) ----------------------------
tmap_save(tmap_arrange_1, 
          "tmap_arrange_1_v5.png", width =6.78,
          height=5, units ='in', asp = 0)
# save 2 columns of tmap_arrange (3 maps) ----------------------------
tmap_save(tmap_arrange_2, 
          "tmap_arrange_2_v5.png", width =6.78,
          height=5, units ='in', asp = 0)

```

## Data management

```{r}

# -------------------------------------------------------------------------------------------------------------
# IMPORTANT TASKS 7-10-20: 
# -------------------------------------------------------------------------------------------------------------
# 1) Confirm that classes and scores were created correctly for indicators and don't require complement values
#    e.g ITN_dist_class (& score)
# COVID_case_class
# Disruptive_event_class (does this require modification or use of mean from UN index?)
# SMC_class (0 or 1) 
# -------------------------------------------------------------------------------------------------------------

# For input chloropleth maps, include the following variables:
# SMC_score, ITN_dist_score, COVID_case_score, Disruptive_event_score,
# temp suitability, rainfall anomaly, built up area/presence index
# slim down for quick map creation
# NOTE: Covid_case_score is being updated by Alyssa on 7-10-20, don't use until updated by you or her
nga_quick_map_slim <- nga_master6[c(1, 14, 17, 19, 21, 33, 34) ] # slim down NGA_master6
View(nga_quick_map_slim)

# Convert nga_7.spdf to dataframe for manipulation
nga_7_v2.df <- as.data.frame(nga_7.spdf)

#Make disruptive_event_class a numeric
nga_7_v2.df$Disruptive_event_class <- as.numeric(nga_7_v2.df$Disruptive_event_class)


```

## Internally displaced population 

```{r}
#---------------------------------------------------------------------
# Create an internally displaced population subset map ---------------
#---------------------------------------------------------------------

idp_map_vars <- c(1, 5, 7, 62, 77, 83 )
idp_map_vars.df <- nga_7_8_29_20.df[idp_map_vars]
View(idp_map_vars.df)

# n = 58. This should consist of the n = 58 LGAs that have
# internally displaced populations
nga_idp <- filter(idp_map_vars.df, idp_eq_wt_score == 0.5)
# what are the malaria risk (factor - variable type) values?
View(nga_idp)

weighted_idp_8_31_20.spdf <- filter(weighted, Disruptive_event_class==1)

# Merge weighted.spdf with nga_idp
# n = 775
weighted_idp.spdf <- merge(adm2.nga, nga_idp, by="GID_2", duplicateGeoms = T)

table(weighted_idp.spdf$Malaria_risk_factor, useNA = "always")
# 1    2    3    4    5 <NA> 
# 4    4   25   18    7    0 
# Note: New results on 8-6-20
# 1    2    3    4 <NA> 
# 4   11   26   17    0 

# Note: New results on 8-31-20
#   1    2    3    4    5    6 <NA> 
# 122  154  185  156  108   50    0

# New results on 9-4-20
# 1    2    3    4    5    6 <NA> 
# 8   30   20    0    0    0  717 

# weighted_idp_2.spdf <- filter(weighted_idp.spdf, Disruptive_event_class == 1)

# get jenks for this subset
getJenksBreaks(weighted_idp.spdf$Malaria_risk_factor, 3, subset = NULL)
# [1] 1 2 3 5
# new jenks run on 8-6-20
# [1] 1 2 3 4
# new jenks run on 9-4-20
# [1] 1 1 2 3
summary(weighted_idp.spdf$Malaria_risk_factor)
# Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
# 1.000   2.250   3.000   2.966   4.000   4.000 

weighted_idp.spdf$idp_risk_group[weighted_idp.spdf$Malaria_risk_factor == 3 ]  <- 3                                        # classify as rank 3
weighted_idp.spdf$idp_risk_group[weighted_idp.spdf$Malaria_risk_factor == 2 ] <- 2  # classify as rank 2
weighted_idp.spdf$idp_risk_group[weighted_idp.spdf$Malaria_risk_factor == 1 ] <- 1                                         # classify as rank 1

# spot check
table(weighted_idp.spdf$idp_risk_group, useNA = "always")
# 1    2    3 <NA> 
# 8   30   20  717  

# convert idp_risk group to factor variable
weighted_idp.spdf$idp_risk_grp_fac <- as.factor(weighted_idp.spdf$idp_risk_group)
# verify same amt of values in groups
table(weighted_idp.spdf$idp_risk_grp_fac, useNA = "always") 
# 1    2    3 <NA> 
# 8   25   25    0 
# new distribution of idp_risk_grp_fac on 8-6-20
#  1    2    3 <NA> 
# 15   26   17    0 

# distrib on 9-4-20
# 1    2    3 <NA> 
# 8   30   20  717 

# Convert weighted_idp.spdf to df
weighted_idp.df <- as.data.frame(weighted_idp.spdf)

# subset nga_idp to make easy to manage 
nga_idp_slim.spdf <- weighted_idp.spdf[c(1, 5, 7, 17,18, 20) ] # slim down nga_idp
# visualize nga_idp_slim
head(nga_idp_slim)

# merge with  adm2.nga to make spdf
nga_7_slim.spdf <- merge(nga_idp_slim, weighted.spdf, by="GID_2", duplicateGeoms = T)

# spot check
table(nga_7_slim.spdf$Malaria_risk_factor, useNA = "always")
#  1    2    3 <NA> 
# 15   26   17    0 
```

## Create internally displaced population risk categories

```{r}
# Create high, med, low risk category variable
nga_7_slim.spdf$idp_risk_level[nga_7_slim.spdf$idp_risk_group == 3] <- "Highest"   # classify as rank 3
nga_7_slim.spdf$idp_risk_level[nga_7_slim.spdf$idp_risk_group == 2] <- "Higher"     # classify as rank 2
nga_7_slim.spdf$idp_risk_level[nga_7_slim.spdf$idp_risk_group == 1] <- "High"  

# Create high, med, low risk category variable
weighted_idp.spdf$idp_risk_level[weighted_idp.spdf$idp_risk_grp_fac == 3] <- "Highest"   # classify as rank 3
weighted_idp.spdf$idp_risk_level[weighted_idp.spdf$idp_risk_grp_fac == 2] <- "Higher"     # classify as rank 2
weighted_idp.spdf$idp_risk_level[weighted_idp.spdf$idp_risk_grp_fac == 1] <- "High"  

# spot check
table(weighted_idp.spdf$idp_risk_level, useNA = "always")
# High  Higher Highest    <NA> 
#   15      26      17       0 

# 9-4-20
# High  Higher Highest    <NA> 
#     8      30      20     717 

# Remove unnecessary variables from table for report
idp_table.spdf <- weighted_idp.spdf[c(1, 5, 7, 17, 20) ] 

idp_table.df <- as.data.frame(idp_table.spdf)

names(idp_table.spdf)

# Rename variables
idp_table.df$LGA <- idp_table.df$NAME_2.x

# View dataset
View(idp_table.df)

# Remove NAME_2.x variable
idp_table_v2.df <- idp_table.df[-c(3)]
idp_table_v2.df <- idp_table_v2.df[c(1, 2, 6, 3, 4, 5) ] 

# View dataset
View(idp_table_v2.df)

# Sort by Risk_category
idp_table_sorted_v2.df <- idp_table_v2.df[order(-idp_table_v2.df$Risk_category.x),]
# View dataset
View(idp_table_sorted_v2.df)

# Export idp subset dataset with risk scores/categories --------
write.csv(idp_table_sorted_v2.df,"/Volumes/Samsung T5/LACIE MacOS Extended/Tulane Research Projects/Malaria Consortium/Nigeria Data/Datasets/idp_table_df_v2.csv", row.names = TRUE)
write.csv(idp_table.df,"/Volumes/Samsung T5/LACIE MacOS Extended/Tulane Research Projects/Malaria Consortium/Nigeria Data/Datasets/idp_table_9_4_20.csv", row.names = TRUE)

# Convert to spdf for map
idp_table_sorted_v3.spdf <- merge(weighted_idp.spdf, idp_table_sorted_v2.df, by="GID_2", duplicateGeoms = T)

```

## Create internally displaced populations map

```{r}

#--------------------------------------------------
# Make IDP LGAs prioritization map ----------------
# n = 58 LGAs -------------------------------------
# -------------------------------------------------

# Assign color palette
PuRd <- get_brewer_pal("PuRd", n = 3)

# Note: Modified on 8-6-20
# assign map to object
idp_58_wghtd_legend_map <- tm_shape(idp_table.spdf) + 
    tm_fill("idp_risk_grp_fac", title = "Malaria risk/\nITN prioritization of\nLGAs with internally\n displaced persons",
            n=2, style = "cat", 
            palette = "PuRd", labels = c("High", "Higher", "Highest"), 
            lwd =0.01, border.col = "black",
            colorNA = NA) + 
    tm_polygons(lwd = 0.5, col = "gray3") +
    tm_scale_bar(position = c(0.65, 0.01)) + 
    tm_layout(frame = FALSE, legend.outside = FALSE,
              legend.position = c(0.75,0.07), legend.title.size = 1) +
    tm_shape(adm1.nga) + tm_borders(col = "grey40", lwd = 0.35)

# Save disruptive event map
# setwd("D:\\LACIE MacOS Extended\\Tulane Research Projects\\Malaria Consortium\\Nigeria Data\\Maps\\disruptive event") #pc
setwd("/Volumes/Samsung T5/LACIE MacOS Extended/Tulane Research Projects/Malaria Consortium/Nigeria Data/Maps/disruptive event") #mac
tmap_save(idp_58_wghtd_legend_map, "idp_58_wghtd_legend_map_9_4_20.png",
          width =6.78, height=5, units ='in', asp = 0)

```


